### 2017-05-12 draft.tex
ok

% \documentclass{llncs}
\documentclass[10pt,conference,letterpaper,twocolumn]{IEEEtran}

\usepackage[usenames,dvipsnames]{color}

\usepackage{latexsym}
\usepackage{amsxtra} 
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{pslatex}
\usepackage{epsfig}
\usepackage{wrapfig}
\usepackage{paralist}
\usepackage{graphics}
\usepackage{stmaryrd}
\usepackage{txfonts}
\usepackage{framed}
\usepackage{makecell}
\usepackage{url}
\usepackage{tikz}
\usetikzlibrary{automata,positioning, calc}
\usepackage[inline,shortlabels]{enumitem}

\usepackage{proof}

\usepackage{algorithm}
\usepackage{algorithmicx}
\usepackage[noend]{algpseudocode}

\usepackage[draft]{commenting}

\pagestyle{plain}

\include{commands}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{document}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\title{The Impact of Alternation}

\author{Radu Iosif and Xiao Xu \\
Verimag/CNRS/Universit\'e de Grenoble Alpes\\
Email: \{Radu.Iosif,Xiao.Xu\}@univ-grenoble-alpes.fr}

\maketitle

\begin{abstract}
Alternating automata have been widely used to model and verify systems
that handle data from finite domains, such as communication protocols
or hardware. The main advantage of the alternating model of
computation is that complementation is possible in linear time, thus
allowing to concisely encode trace inclusion problems that occur often
in verification. In this paper we consider alternating automata over
infinite alphabets, whose transition rules are formulae in a combined
theory of booleans and some infinite data domain, that relate past and
current values of the data variables. The data theory is not fixed,
but rather it is a parameter of the class. We show that union,
intersection and complementation are possible in linear time in this
model and, though the emptiness problem is undecidable, we provide two
efficient semi-algorithms, inspired by two state-of-the-art
abstraction refinement model checking methods: lazy predicate
abstraction \cite{HJMS02} and the \impact~ semi-algorithm
\cite{mcmillan06}. We have implemented both methods and report the
results of an experimental comparison.
\end{abstract}

\section{Introduction}

\section{Preliminaries}

A \emph{signature} $\sig = (\ssorts{\sig},\sfuns{\sig})$ consists of a
set $\ssorts{\sig}$ of \emph{sort symbols} and a set $\sfuns{\sig}$ of
sorted \emph{function symbols}. To simplify the presentation, we
assume w.l.o.g. that $\ssorts{\sig} = \set{\Data,\Bool}$\footnote{The
  generalization to more than two sorts is without difficulty, but
  would unnecessarily clutter the technical presentation.} and each
function symbol $f \in \sfuns{\sig}$ has $\#(f) \geq 0$ arguments of
sort $\Data$ and return value $\sigma(f) \in \ssorts{\sig}$. If
$\#(f)=0$ then $f$ is a \emph{constant}. We consider the constants
$\top$ and $\bot$ of sort $\Bool$.

We consider an infinite countable set of \emph{variables} $\vars$,
where each $x \in \vars$ has an associated sort $\sigma(x)$. Terms are
defined recursively: a term $t$ of sort $\sigma(t)=S$ is a variable $x
\in \vars$ where $\sigma(x)=S$, or $f(t_1,\ldots,t_{\#(f)})$ where
$t_1,\ldots,t_{\#(f)}$ are terms of sort $\Data$ and $\sigma(f)=S$. An
\emph{atom} is a term of sort $\Bool$ or an equality $t \teq s$
between two terms of sort $\Data$. A \emph{literal} is an atom or its
negation. A \emph{formula} is an existentially quantified combination
of atoms using disjunction $\vee$, conjunction $\wedge$ and negation
$\neg$ and we write $\phi \rightarrow \psi$ for $\neg\phi \vee \psi$.

We denote by $\fv{\sigma}{\phi}$ the set of free (not under the scope
of an existential quantifier) variables of sort $\sigma$ in $\phi$ and
write $\fv{}{\phi}$ for $\bigcup_{\sigma \in \ssorts{\sig}}
\fv{\sigma}{\phi}$. For a variable $x \in \fv{}{\phi}$ and a term $t$
such that $\sigma(t) = \sigma(x)$, let $\phi[t/x]$ be the result of
replacing each occurrence of $x$ by $t$. For indexed sets
$\vec{t}=\set{t_1,\ldots,t_n}$ and $\vec{x}=\set{x_1,\ldots,x_n}$, we
write $\phi[\vec{t}/\vec{x}]$ for the formula obtained by
simultaneously replacing $x_i$ with $t_i$ in $\phi$, for all
$i\in[1,n]$. The size $\len{\phi}$ is the number of symbols occuring
in $\phi$.

An \emph{interpretation} $\I$ maps\begin{inparaenum}[(1)]
\item the sort $\Data$ into a non-empty set $\Data^\I$, 
%
\item the sort $\Bool$ into the set $\booli = \set{\true,\false}$, where
  $\top^\I = \true$, $\bot^\I = \false$, and
%
\item each function symbol $f$ into a total function $f^\I :
  (\Data^\I)^{\#(f)} \rightarrow \sigma(f)^I$, or an element of
  $\sigma(f)^I$ when $\#(f)=0$.
\end{inparaenum}
Given an interpretation $\I$, a \emph{valuation} $\nu$ maps each
variable $x \in \vars$ into an element $\nu(x) \in \sigma(x)^\I$. For
a term $t$, we denote by $t^\I_\nu$ the value obtained by replacing
each function symbol $f$ by its interpretation $f^\I$ and each
variable $x$ by its valuation $\nu(x)$. For a formula $\phi$, we write
$\I,\nu \models \phi$ if the formula obtained by replacing each term
$t$ in $\phi$ by the value $t^\I_\nu$ is logically equivalent to true.
A formula $\phi$ is \emph{satisfiable} in the interpretation $\I$ if
there exists a valuation $\nu$ such that $\I,\nu \models \phi$, and
\emph{valid} if $\I,\nu \models \phi$ for all valuations $\nu$.  The
\emph{theory} $\theory(\sig,\I)$ is the set of valid formulae written
in the signature $\sig$, with the interpretation $\I$. A decision
procedure for $\theory(\sig,\I)$ is an algorithm that takes a formula
$\phi$ in the signature $\sig$ and returns yes iff $\phi \in
\theory(\sig,\I)$.

Given formulae $\varphi$ and $\psi$, we say that \emph{$\phi$ entails
  $\psi$}, denoted $\phi \models^\I \psi$ iff $\I,\nu \models \varphi$
implies $\I,\nu \models \psi$, for each valuation $\nu$, and $\phi
\iffeq^\I \psi$ iff $\phi \models^\I \psi$ and $\psi \models^\I \phi$.
We omit mentioning the interpretation $\I$ when it is clear from the
context.

\section{Alternating Data Automata}

In the rest of this section we fix an interpretation $\I$ and a finite
alphabet $\Sigma$ of \emph{input events}. Given a finite set $\vec{x}
\subset \vars$ of variables of sort $\Data$, let $\vec{x} \mapsto
\Data^\I$ be the set of valuations of the variables $\vec{x}$ and
$\Sigma[\vec{x}] = \Sigma \times (\vec{x} \mapsto \Data^\I)$ be the
set of \emph{data symbols}. A \emph{data word} (word in the sequel) is
a finite sequence $(a_1,\nu_1)(a_2,\nu_2) \ldots (a_n,\nu_n)$ of data
symbols, where $a_1,\ldots,a_n \in \Sigma$ and $\nu_1,\ldots,\nu_n :
\vec{x} \rightarrow \Data^\I$ are valuations. We denote by
$\varepsilon$ the empty sequence, by $\Sigma^*$ the set of finite
sequences of input events and by $\Sigma[\vec{x}]^*$ the set of data
words over $\vec{x}$.

This definition generalizes the classical notion of words from a
finite alphabet to the possibly infinite alphabet
$\Sigma[\vec{x}]$. Clearly, when $\Data^\I$ is sufficiently large or
infinite, we can map the elements of $\Sigma$ into designated elements
of $\Data^\I$ and use a special variable to encode the input
events. However, keeping $\Sigma$ explicit in the following simplifies
several technical points below, without cluttering the presentation.

Given sets of variables $\vec{b},\vec{x} \subset \vars$ of sort
$\Bool$ and $\Data$, respectively, we denote by
$\Form(\vec{b},\vec{x})$ the set of formulae $\phi$ such that
$\fv{\Bool}{\phi} \subseteq \vec{b}$ and $\fv{\Data}{\phi} \subseteq
\vec{x}$. By $\Form^+(\vec{b},\vec{x})$ we denote the set of formulae
from $\Form(\vec{b},\vec{x})$ in which each boolean variable occurs
under an even number of negations.

An \emph{alternating data automaton} (ADA or automaton in the sequel)
is a tuple $\A = \tuple{\vec{x},Q,\iota,F,\Delta}$,
where: \begin{compactitem}
%
\item $\vec{x} \subset \vars$ is a finite set of variables of sort
  $\Data$,
%
\item $Q$ is a finite set of variables of sort $\Bool$, called
  \emph{states},
%
\item $\iota \in \Form^+(Q,\emptyset)$ is positive boolean combination
  of states defining the \emph{initial configurations},
%
\item $F \subseteq Q$ is a set of \emph{final states}, and
%
\item $\Delta : Q \times \Sigma \rightarrow
  \Form^+(Q,\overline{\vec{x}}\cup\vec{x})$ is a \emph{transition
    function}, where $\overline{\vec{x}}$ is the set $\{\overline{x}
  \mid x \in \vec{x}\}$.
\end{compactitem}
Intuitively, in each formula $\Delta(q,a)$ describing a transition
rule, the variables $\overline{\vec{x}}$ track the previous and
$\vec{x}$ the current values of the variables of $\A$. Observe
further, that the initial values of the variables are left
unconstrained, as the initial configuration does not contain free
data variables. The size of $\A$ is defined as $\len{\A} = \len{\iota}
+ \sum_{(q,a) \in Q\times\Sigma}\len{\Delta(q,a)}$.

Formally, let $\vec{x}_k = \set{x_k \mid x \in \vec{x}}$, $k\geq0$, be
a set of time-stamped variables. For an input event $a \in \Sigma$ and
a formula $\phi$, we write $\Delta(\phi,a)$ (respectively,
$\Delta^k(\phi,a)$) for the formula obtained from $\phi$ by
simultaneously replacing each state $q \in \fv{\Bool}{\phi}$ by the
formula $\Delta(q,a)$ (respectively
$\Delta(q,a)[\vec{x}_k/\overline{\vec{x}},\vec{x}_{k+1}/\vec{x}]$, for
all $k\geq0$). Given a word $w = (a_1,\nu_1)(a_2,\nu_2) \ldots
(a_n,\nu_n)$, the \emph{run} of $\A$ over $w$ is the sequence of
formulae $\phi_0,\phi_1,\ldots,\phi_n$, where $\phi_0 \equiv \iota$
and, for all $k\in[1,n]$, we have $\phi_k \equiv
\Delta^k(\phi_{k-1},a_k)$. Next, we slightly abuse notation and write
$\Delta(\iota,a_1,\ldots,a_n)$ for the formula
$\phi_n(\vec{x}_0,\ldots,\vec{x}_n)$ above. We say that $\A$
\emph{accepts} $w$ iff $\I,\nu \models \Delta(\iota,a_1,\ldots,a_n)$,
where $\nu$ is some valuation that maps:\begin{inparaenum}[(1)]
%
\item each $x \in \vec{x}_k$ to $\nu_k(x)$, for all $k\in[1,n]$, 
%
\item each $q \in \fv{\Bool}{\phi_n} \cap F$ to $\top$ and 
%
\item each $q \in \fv{\Bool}{\phi_n} \setminus F$ to $\bot$.
\end{inparaenum}
The language of $\A$ is the set $L(\A)$ of words from
$\Sigma[\vec{x}]^*$ accepted by $\A$.
%
In this paper we tackle the following problems: \begin{compactenum}
\item \emph{boolean closure}: given automata $\A_1$ and $\A_2$, both
  with the same set of variables $\vec{x}$, do there exist automata
  $\A_\cup$, $\A_\cap$ and $\overline{\A_1}$ such that $L(\A_\cup) =
  \A_1 \cup \A_2$, $L(A_\cap) = \A_1 \cap \A_2$ and
  $L(\overline{\A_1}) = \Sigma[\vec{x}]^* \setminus L(\A_1)$ ?
%
\item \emph{emptiness}: given an automaton $\A$, is it the case that
  $L(\A) = \emptyset$ ?
\end{compactenum}

It is well known that other problems, such as \emph{universality}
(given automaton $\A$ with variables $\vec{x}$, does $L(\A) =
\Sigma[\vec{x}]^*$?) and \emph{inclusion} (given automata $\A_1$ and
$\A_2$ with the same set of variables, does $L(\A_1) \subseteq
L(\A_2)$?) can be reduced to the above problems. Observe furthermore
that we do not consider cases in which the sets of variables in the
two automata differ. An interesting problem in this case would be: given
automata $\A_1$ and $\A_2$, with variables $\vec{x}_1$ and
$\vec{x}_2$, respectively, such that $\vec{x}_1 \subseteq \vec{x}_2$,
does $L(\A_1) \subseteq \proj{L(\A_2)}{\vec{x}_1}$, where
$\proj{L(\A_2)}{\vec{x}_1}$ is the projection of the set of words
$L(\A_2)$ onto the variables $\vec{x}_1$? This problem is considered
as future work.

\subsection{Boolean Closure}

Given a set $Q$ of boolean variables and a set $\vec{x}$ of variables
of sort $\Data$, for a formula $\phi \in \Form^+(Q,\vec{x})$, with no
negated occurrences of the boolean variables, we define the formula
$\overline{\phi} \in \Form^+(Q,\vec{x})$ recursively on the structure
of $\phi$:
\[\begin{array}{lclclcl}
\overline{\phi_1 \vee \phi_2} & \equiv & \overline{\phi_1} \wedge \overline{\phi_2} && 
\overline{\phi_1 \wedge \phi_2} & \equiv & \overline{\phi_1} \vee \overline{\phi_2} \\ 
\overline{\neg\phi} & \equiv & \neg \overline{\phi} \text{ if $\phi$ not atom} &&
\overline{\phi} & \equiv & \phi \text{ if $\phi \in Q$} \\
\overline{\phi} & \equiv & \neg\phi \text{ if $\phi \not\in Q$ atom}
\end{array}\]
It is easy to check that $\len{\overline{\phi}} = \len{\phi}$, for
every formula $\phi \in \Form^+(Q,\vec{x})$. If $\phi$ does not
contain boolean variables, then $\overline{\phi}$ and $\neg\phi$ are
equivalent, and in general:

\begin{proposition}\label{prop:overline}
  Given a formula $\phi \in \Form^+(Q,\vec{x})$ and a valuation $\nu$
  mapping each $q \in Q$ to a value $\nu(q) \in \booli$ and each $x \in
  \vec{x}$ to a value $\nu(x) \in \Data^\I$, let $\nu'$ be the
  valuation that assigns each $q \in Q$ the value $\neg\nu(q)$ and
  each $x \in \vec{x}$ the value $\nu(x)$. Then we have 
  $\I,\nu \models \phi$ if and only if $\I,\nu' \not\models \overline{\phi}$. 
\end{proposition}
\proof{Immediate, by induction on the structure of $\phi$. \qed}

In the following let $\A_i =
\tuple{\vec{x},Q_i,\iota_i,F_i,\Delta_i}$, for $i=1,2$, where
w.l.o.g. we assume that $Q_1 \cap Q_2 = \emptyset$. We define: 
\[\begin{array}{rcl}
\A_\cup & = & \tuple{\vec{x},Q_1\cup Q_2,\iota_1 \vee \iota_2,F_1\cup F_2,\Delta_1\cup\Delta_2} \\ 
\A_\cap & = & \tuple{\vec{x},Q_1\cup Q_2,\iota_1 \wedge \iota_2, F_1\cup F_2, \Delta_1\cup\Delta_2} \\ 
\overline{\A_1} & = & \langle \vec{x},Q_1,\iota_1,Q_1\setminus F_1,\overline{\Delta_1} \rangle
\end{array}\] 
where $\overline{\Delta_1}(q,a) \equiv \overline{\Delta_1(q,a)}$, for
all $q \in Q_1$ and $a \in \Sigma$. The following lemma shows the
correctness of the above definitions:

\begin{lemma}\label{lemma:closure}
  Given automata $\A_i = \tuple{\vec{x},Q_i,\iota_i,F_i,\Delta_i}$,
  for $i=1,2$, such that $Q_1 \cap Q_2 = \emptyset$, we have
  $L(\A_\cup) = L(\A_1) \cup L(\A_2)$, $L(\A_\cap) = L(\A_1) \cap
  L(\A_2)$ and $L(\overline{\A_1}) = \Sigma[\vec{x}]^* \setminus
  L(\A_1)$.
\end{lemma}
\proof{ We prove $L(\A_\cup) = L(\A_1) \cup L(\A_2)$ first, the proof
  for $\A_\cap$ being analogous. Let $w = (a_1,\nu_1) \ldots
  (a_n,\nu_n)$ be a word, where $n=0$ corresponds to the empty
  word. We prove by induction on $n\geq0$ that $\Delta(\iota_1 \vee
  \iota_2,a_1\ldots a_n) \iffeq \Delta(\iota_1,a_1\ldots a_n) \vee
  \Delta(\iota_2,a_1\ldots a_n)$. The case $n=0$ follows from the
  definition of the initial configuration of $\A_\cup$. For the
  inductive step $n>0$, $\Delta(\iota_1 \vee \iota_2,a_1\ldots a_n)$
  is obtained from $\Delta(\iota_1 \vee \iota_2,a_1\ldots a_{n-1})$ by
  replacing each variable $q \in \fv{\Bool}{\Delta(\iota_1 \vee
    \iota_2,a_1\ldots a_{n-1})}$ with
  $\Delta(q,a_n)[\vec{x}_{n-1}/\overline{\vec{x}},\vec{x}_n/\vec{x}]$,
  denoted $\Delta^n(\Delta(\iota_1 \vee \iota_2,a_1\ldots
  a_{n-1}),a_n)$. Since by induction hypothesis, $\Delta(\iota_1 \vee
  \iota_2,a_1\ldots a_{n-1}) \iffeq \Delta(\iota_1,a_1\ldots a_{n-1})
  \vee \Delta(\iota_2,a_1\ldots a_{n-1})$, we obtain:
  \[\begin{array}{ll}
  \Delta^n(\Delta(\iota_1 \vee \iota_2,a_1\ldots a_{n-1}),a_n) & \iffeq \\
  \Delta^n(\Delta(\iota_1,a_1\ldots a_{n-1}),a_n) \vee \Delta^n(\Delta(\iota_2,a_1\ldots a_{n-1}),a_n) & \iffeq \\
  \Delta(\iota_1,a_1\ldots a_{n}) \vee \Delta(\iota_2,a_1\ldots a_{n}) \enspace.
  \end{array}\] 
  To prove $L(\overline{\A_1}) = \Sigma[\vec{x}]^* \setminus L(\A_1)$,
  let $w = (a_1,\nu_1) \ldots (a_n,\nu_n)$ be a word and show that
  $\overline{\Delta}(\iota_1,a_1 \ldots a_n) =
  \overline{\Delta(\iota_1,a_1 \ldots a_n)}$ by induction on
  $n\geq0$. The case $n=0$ is immediate, because $\fv{}{\iota_1}
  \subseteq Q$ and thus $\overline{\iota_1} \equiv \iota_1$. For the case
  $n>0$, we compute:
  \[\begin{array}{lcl}
  \overline{\Delta}^n(\overline{\Delta}(\iota_1,a_1\ldots a_{n-1}),a_n) & \iffeq & \text{ (by ind. hyp.)} \\
  \overline{\Delta}^n(\overline{\Delta(\iota_1,a_1\ldots a_{n-1})},a_n) & \iffeq & \text{ (by the def. of $\overline{\phi}$)} \\
  \overline{\Delta(\iota_1,a_1\ldots,a_n)}
  \end{array}\]
  Let $\nu,\nu' : (Q \cup \bigcup_{i=0}^n\vec{x}_i) \rightarrow (\booli
  \cup \Data^\I)$ be valuations such that: \begin{compactitem}
  \item $\nu(q) = \top$ and $\nu'(q) = \bot$, for each $q \in F$, 
    %
  \item $\nu(q) = \bot$ and $\nu'(q) = \top$, for each $q \in Q \setminus F$, 
    %
  \item $\nu(x) = \nu'(x)$, for each $x \in \vec{x}_0$, 
    %
  \item $\nu(x) = \nu'(x) = \nu_i(x)$, for each $x \in \vec{x}_i$ and each $i
    \in [1,n]$.
  \end{compactitem}
  By Proposition \ref{prop:overline}, we have $\I,\nu \models
  \Delta(\iota_1,a_1 \ldots a_n) \iff \I,\nu' \not\models
  \overline{\Delta(\iota_1,a_1 \ldots a_n)} \iff \I,\nu' \not\models
  \overline{\Delta}(\iota_1,a_1 \ldots a_n)$. Thus for all $w \in
  \Sigma[\vec{x}]^*$, we have $w \in L(\A_1) \iff w \not\in
  L(\overline{\A_1})$, i.e. $L(\overline{\A_1}) = \Sigma[\vec{x}]^*
  \setminus L(\A_1)$. \qed}

It is easy to see that $\len{\A_\cup} = \len{\A_\cap} = \len{\A_1} +
\len{\A_2}$ and $\len{\overline{\A}} = \len{\A}$, thus the automata
for the boolean operations, including complementation, can be built in
linear time. This is not surprising, as linear time intersection and
complementation are also possible for alternating automata over finite
alphabets \cite{ChandraKozenStockmeyer81}.

\section{Antichains and Interpolants for Emptiness}

Unlike the boolean closure properties, showed to be effectively
decidable (Lemma \ref{lemma:closure}), the emptiness problem is
undecidable, even in very simple cases. For instance, if $\Data^\I$ is
the set of positive integers, an ADA can simulate an Alternating
Vector Addition System with States (AVASS) \cite{LazicSchmitz14} using
only atoms $x \geq k$ and $x = \overline{x} + k$, for $k \in \zed$,
with the classical interpretation of the function symbols on
integers. Since reachability of a control state is undecidable for
AVASS \cite{LINCOLN92}, the emptiness problem is undecidable for ADA.

Consequently, we give up on the guarantee for termination and build
semi-algorithms that meet the requirements below: \begin{compactenum}[(i)]
\item given an automaton $\A$, if $L(\A) \neq\emptyset$, the procedure
  will terminate and return a word $w \in L(\A)$, and
%
\item if the procedure terminates without returning such a word, then
  $L(\A) = \emptyset$.
\end{compactenum}

Let us fix an automaton $\A = \tuple{\vec{x},Q,\iota,F,\Delta}$ whose
(finite) input event alphabet is $\Sigma$, for the rest of this
section. Given a formula $\phi \in \Form^+(Q,\vec{x})$ and an input
event $a \in \Sigma$, we define the \emph{post-image} function
$\Post{\A}(\phi,a) \equiv \exists \overline{\vec{x}} ~.~
\Delta(\phi[\overline{\vec{x}}/\vec{x}], a) \in \Form^+(Q,\vec{x})$,
mapping each formula in $\Form^+(Q,\vec{x})$ to a formula defining the
effect of reading the event $a$.

We generalize the post-image function to finite sequences of input
events, as follows, for any $u \in \Sigma^*$:
\[\begin{array}{l}
\Post{\A}(\phi,\varepsilon) \equiv \phi \hspace*{5mm} \Post{\A}(\phi,ua) \equiv \Post{\A}(\Post{\A}(\phi,u),a) \\
\Accept{\A}(u) \equiv \Post{\A}(\iota,u) \wedge \bigwedge_{q \in Q \setminus F} (q \rightarrow \bot)
\end{array}\]
Then the emptiness problem for $\A$ becomes: does there exist $u \in
\Sigma^*$ such that the formula $\Accept{\A}(u)$ is satisfiable?
Observe that, since we ask a satisfiability query, the final states of
$\A$ need not be constrained\footnote{ Since each state occurs
  positively in $\Accept{\A}(u)$, this formula has a model iff it has
  a model with every $q \in F$ set to true.}. A na\"ive semi-algorithm
enumerates all finite sequences and checks the satisfiability of
$\Accept{\A}(u)$ for each $u \in \Sigma^*$, using a decision procedure
for the theory $\theory(\sig,\I)$.

Since no boolean variable from $Q$ occurs under negation in $\phi$, it
is easy to prove the following monotonicity property: given two
formulae $\phi,\psi \in \Form^+(Q,\vec{x})$ if $\phi \models \psi$
then $\Post{\A}(\phi,u) \models \Post{\A}(\psi,u)$, for any $u \in
\Sigma^*$. This suggest an improvement of the above semi-algorithm,
that enumerates and stores only a set $U \subseteq \Sigma^*$ for which
$\set{\Post{\A}(\phi,u) \mid u \in U}$ forms an
\emph{antichain}\footnote{Given a partial order $(D,\preceq)$ an
  antichain is a set $A \subseteq D$ such that $a \not\preceq b$ for
  any $a,b \in A$.} w.r.t. the entailment partial order. This is
because, for any $u,v \in \Sigma^*$, if $\Post{\A}(\iota,u) \models
\Post{\A}(\iota,v)$ and $\Accept{\A}(uw)$ is satisfiable for some $w
\in \Sigma^*$, then $\Post{\A}(\iota,uw) \models \Post{\A}(\iota,vw)$,
thus $\Accept{\A}(vw)$ is satisfiable as well, and there is no need
for $u$, since the non-emptiness of $\A$ can be proved using $v$
alone. However, even with this optimization, the enumeration of
sequences from $\Sigma^*$ diverges in many real cases, because
infinite antichains exist in many interpretations, e.g.\ $q \wedge x
\teq 0,~ q \wedge x \teq 1, \ldots$ for $\Data^\I = \nat$.

A \emph{safety invariant} for $\A$ is a function $\Inv : (Q \mapsto
\booli) \rightarrow 2^{\vec{x} \mapsto \Data^\I}$ such that, for every
boolean valuation $\beta : Q \rightarrow \booli$, every valuation $\nu
: \vec{x} \mapsto \Data^\I$ of the data variables and every finite
sequence $u \in \Sigma^*$ of input events, the following
hold: \begin{compactenum}
%% \item $\beta \models \iota \Rightarrow \nu_0 \in \Inv(\beta)$, where
%%   $\nu_0(x) = \mathbf{0}$, for all $x \in \vec{x}$.
%
\item\label{it1:inv} $\I,\beta \cup \nu \models \Post{\A}(\iota,u)
  \Rightarrow \nu \in \Inv(\beta)$, and
%
\item\label{it2:inv} $\nu \in \Inv(\beta) \Rightarrow \I,\beta \cup
  \nu \not\models \Accept{\A}(u)$.
\end{compactenum}
If $\Inv$ satisfies only the first point above, we call it an
\emph{invariant}. Intuitively, a safety invariant maps every boolean
valuation into a set of data valuations, that contains the initial
configuration $\iota = \Post{\A}(\iota,\varepsilon)$, whose data
variables are unconstrained, over-approximates the set of reachable
valuations (point \ref{it1:inv}) and excludes the valuations satisfying the
acceptance condition (point \ref{it2:inv}).  A formula $\phi(Q,\vec{x})$
\emph{defines} $\Inv$ iff for all $\beta: Q \rightarrow \booli$ and
$\nu : \vec{x} \rightarrow \Data^\I$, we have $\I,\beta\cup\nu \models
\phi$ iff $\nu \in \Inv(\beta)$.

\begin{lemma}\label{lemma:safety-invariant}
  For any automaton $\A$, we have $L(\A) = \emptyset$ if and only if
  $\A$ has a safety invariant.
\end{lemma}
\proof{ Let $\A = \tuple{\vec{x},Q,\iota,F,\Delta}$ in the following.
  ``$\Leftarrow$'' This direction is trivial. ``$\Rightarrow$'' We
  define $\Inv : (Q \mapsto \booli) \rightarrow 2^{\vec{x} \mapsto
    \Data^\I}$ as follows. For each $\beta : Q \rightarrow \booli$,
  let $\Inv(\beta) = \{\nu : \vec{x} \rightarrow \Data^\I \mid \exists
  u \in \Sigma^* ~.~ \beta\cup\nu \models
  \Post{\A}(\iota,u)\}$. Checking that $\Inv$ is a safety invariant is
  straightforward. \qed}

Turning back to the issue of divergence of language emptiness
semi-algorithms in the case $L(\A) = \emptyset$, we can observe that
an enumeration of input sequences $u_1,u_2,\ldots \in \Sigma^*$ can
stop at step $k$ as soon as $\bigvee_{i=1}^k \Post{\A}(\iota,u_i)$
defines a safety invariant for $\A$. Although this condition can be
effectively checked using a decision procedure for the theory
$\theory(\sig,\I)$, there is no guarantee that this check will ever
succeed.

The solution we adopt in the sequel is abstraction to ensure the
termination of invariant computations. However, it is worth pointing
out from the start that abstraction alone will only allow us to build
invariants that are not necessarily safety invariants. To meet the
latter condition, we resort to counterexample guided abstraction
refinement (CEGAR).

Formally, we fix a set of formulae $\preds \subseteq
\Form(Q,\vec{x})$, such that $\bot \in \preds$ and refer to these
formulae as \emph{predicates}. Given a formula $\phi$, we denote by
$\abs{\phi} \equiv \bigwedge \set{\pi \in \preds \mid \phi \models
  \pi}$ the abstraction of $\phi$ w.r.t. the predicates in
$\preds$. The abstract versions of the post-image and acceptance
condition are defined as follows, for any $u \in \Sigma^*$:
\[\begin{array}{l}
\AbsPost{\A}(\phi,\varepsilon) \equiv \phi \hspace*{5mm}
\AbsPost{\A}(\phi,ua) \equiv \abs{(\Post{\A}(\AbsPost{\A}(\phi,u),a))} \\
\AbsAccept{\A}(u) \equiv \AbsPost{\A}(\iota,u) \wedge \bigwedge_{q \in Q \setminus F} 
(q \rightarrow \bot)
\end{array}\]
\begin{lemma}\label{lemma:abstract-invariant}
  For any bijection $\mu : \nat \rightarrow \Sigma^*$, there exists
  $k>0$ such that $\bigvee_{i=0}^k \AbsPost{\A}(\iota,\mu(i))$ defines
  an invariant for $\A$. 
\end{lemma}
\proof{ It is sufficient to show that there exists $k\geq0$ such that for
  all $u \in \Sigma^*$ there exists $i \in [0,k]$ such that
  $\Post{\A}(\iota,u) \models \AbsPost{\A}(\iota,\mu(i))$. We have
  $\Post{\A}(\iota,u) \models \AbsPost{\A}(\iota,u)$ for all $u \in
  \Sigma^*$. But since $\preds$ is a finite set, also the set
  $\{\AbsPost{\A}(\iota,u) \mid u \in \Sigma^*\}$ is finite. Thus
  there exists $k \geq 0$ such that, for all $u \in \Sigma^*$ there
  exists $i \in [0,k]$ such that $\AbsPost{\A}(\iota,u) \iffeq
  \AbsPost{\A}(\iota,\mu(i))$, which concludes the proof. \qed}

\paragraph{\em Remark}
While computing the concrete post-image $\Post{\A}$ relies on an
expensive quantifier elimination, it is common belief that this should
hold also for $\AbsPost{\A}$. However, the choice of conjoined
predicates that form $\AbsPost{\A}$ requires only a quantifier-free
satisfiability query: $\Post{\A}(\phi,a) \models \pi \iff \exists
\overline{\vec{x}} ~.~ \Delta(\phi[\overline{\vec{x}}/\vec{x}],a)
\models \pi \iff \Delta(\phi[\overline{\vec{x}}/\vec{x}],a) \wedge
\neg\pi$ unsatisfiable. \qed

We are left with fulfilling point (\ref{it2:inv}) from the definition
of a safety invariant. To this end, suppose that, for a given set
$\preds$ of predicates, the invariant $\abs{\Inv}$, defined by the
previous lemma, meets point (\ref{it1:inv}) but not point
(\ref{it2:inv}), where $\Post{\A}$ and $\Accept{\A}$ replace
$\AbsPost{\A}$ and $\AbsAccept{\A}$, respectively. In other words,
there exists a finite sequence $u \in \Sigma^*$ such that $\nu \in
\abs{\Inv}(\beta)$ and $\I,\beta\cup\nu \models \AbsAccept{\A}(u)$,
for some boolean $\beta : Q \rightarrow \booli$ and data $\nu :
\vec{x} \rightarrow \Data^\I$ valuations. Such a $u\in\Sigma^*$ is
called a \emph{counterexample}.

Once a counterexample $u$ is discovered, there are two
possibilities. Either\begin{inparaenum}[(i)]
\item $\Accept{\A}(u)$ is satisfiable, in which case $u$ is
  \emph{feasible} and $L(\A) \neq \emptyset$, or
%
\item $\Accept{\A}(u)$ is unsatisfiable, in which case $u$ is
  \emph{spurious}.
\end{inparaenum}
In the first case, our semi-algorithm stops and returns a witness for
non-emptiness, obtained from the satisfying valuation of
$\Accept{\A}(u)$ and in the second case, we must strenghten the
invariant by excluding from $\abs{\Inv}$ all pairs $(\beta,\nu)$ such
that $\I,\beta\cup\nu \models \AbsAccept{\A}(u)$. This strenghtening
is carried out by adding to $\preds$ several predicates that are
sufficient to exclude the spurious counterexample.

In general, given an unsatisfiable conjunction $\Phi \equiv
\phi_1(X_0,X_1) \wedge \phi_2(X_1,X_2) \wedge \ldots \wedge
\phi_n(X_{n-1},X_n)$ of time-stamped variables $X_i = \set{x_i \mid x
  \in X}$ of any sort, a solution of the \emph{interpolation problem}
$\Phi$, simply called an \emph{interpolant}, is a tuple $\tuple{I_0(X),
  I_1(X), \ldots, I_n(X)}$ such that:\begin{inparaenum}[(i)]
\item $I_0 \equiv \top$,  
%
\item $I_{i-1}[X_{i-1}/X] \wedge \phi_i(X_{i-1},X_i) \models
  I_i[X_i/X]$, for all $i \in [1,n]$, and
%
\item $I_n \equiv \bot$. 
\end{inparaenum}
In the following, we shall assume the existence of an interpolating
decision procedure for $\theory(\sig,\I)$.

\paragraph{\em Remark} 
We deviate from the standard notion of \emph{generalized interpolants}
\cite{mcmillan06}, which requires $\fv{}{I_i} \subseteq \fv{}{\phi_i}
\cap \fv{}{\phi_{i+1}}$, for each $i \in [1,n-1]$, and assume further
that the variables have been already renamed by removing the
time-stamps. \qed

A classical method for abstraction refinement is to add the elements
of the interpolant obtained from a proof of spuriousness to the set of
predicates. This guarantees progress, meaning that the particular
spurious counterexample, from which the interpolant was generated,
will never be revisited in the future. Though not always, in many
practical test cases this progress property eventually yields a safety
invariant.

Given a non-empty spurious counterexample $u = a_1\ldots a_n$, where
$n>0$, we consider the following interpolation problem: 
\begin{eqnarray}\label{eq:interpolation-problem}
\Theta(u) & \equiv & \theta_0(Q_0) \wedge \theta_1(Q_0 \cup Q_1,\vec{x}_0
\cup \vec{x}_1) \wedge \ldots \\ 
&& \wedge~ \theta_n(Q_{n-1} \cup Q_n,\vec{x}_{n-1} \cup \vec{x}_n) \wedge \theta_{n+1}(Q_n) \nonumber
\end{eqnarray}
where $Q_k = \set{q_k \mid q \in Q}$, $k \in [0,n]$ are time-stamped
sets of boolean variables corresponding to the set $Q$ of states of
$\A$. The first conjunct $\theta_0(Q_0) \equiv \iota[Q_0/Q]$ is the
initial configuration of $\A$, with every $q \in \fv{\Bool}{\iota}$
replaced by $q_0$. The definition of $\theta_k$, $k\in[1,n]$ uses
\emph{replacement sets} $R_\ell \subseteq Q_\ell$, $\ell\in [0,n]$,
which are defined inductively below: \begin{compactitem}
\item $R_0 = \fv{\Bool}{\theta_0}$ and 
%
\item $\theta_\ell \equiv \bigwedge_{q_{\ell-1}\in R_{\ell-1}}
  (q_{\ell-1} \rightarrow
  \Delta(q,a_\ell)[Q_\ell/Q,\vec{x}_{\ell-1}/\overline{\vec{x}},\vec{x}_\ell/\vec{x}])$
  and $R_\ell = \fv{\Bool}{\theta_\ell} \cap Q_\ell$, for each
  $\ell\in[1,n]$.
\end{compactitem}
Finally, we set $\theta_{n+1}(Q_n) \equiv \bigwedge_{q \in Q \setminus
  F} (q_n \rightarrow \bot)$. The intuition is that $\theta_0, \ldots,
\theta_n$ encode the run of $\A$ over $u$ and $\theta_{n+1}$ is the
acceptance condition, which forces the last remaining non-final states
to be false.

\begin{proposition}\label{prop:asterisk}
  Given a formula $\phi \in \Form^+(Q,\vec{x})$ and $a \in \Sigma$, we
  have \(\Delta(\phi,a) \iffeq \exists Q' ~.~ \phi[Q'/Q] \wedge
  \bigwedge_{q \in Q} (q' \rightarrow \Delta(q,a))\).
\end{proposition}
\proof{``$\Rightarrow$'' If $\I,\beta\cup\overline\nu\cup\nu \models
  \Delta(\phi,a)$, for some valuations $\beta : Q \rightarrow \booli$
  and $\overline\nu : \overline{\vec{x}} \rightarrow \Data^\I$, $\nu :
  \vec{x} \rightarrow \Data^\I$, then we build a valuation $\beta' :
  Q' \rightarrow \booli$ such that
  $\I,\beta'\cup\beta\cup\overline\nu\cup\nu \models \phi[Q'/Q] \wedge
  \bigwedge_{q \in Q} (q' \rightarrow \Delta(q,a))$. For each
  occurrence of a formula $\Delta(q,a)$ in $\Delta(\phi,a)$ we set
  $\beta'(q') = \true$ if $\I,\beta\cup\overline\nu\cup\nu \models
  \Delta(q,a)$ and $\beta'(q') = \false$, otherwise. Since there are
  no negated occurrences of such subformulae, the definition of
  $\beta'$ is consistent, and the check
  $\I,\beta'\cup\beta\cup\overline\nu\cup\nu \models \phi[Q'/Q] \wedge
  \bigwedge_{q \in Q} (q' \rightarrow \Delta(q,a))$ is
  immediate. $''\Leftarrow''$ This direction is an easy check. \qed}

Moreover, we require that an interpolant
$\tuple{\top,I_0,\ldots,I_n,\bot}$ for the interpolation problem
$\Theta(u)$ does not have negative occurrences of states, i.e.\ $I_i
\in \Form^+(Q,\vec{x})$, for all $i \in [0,n]$. Such an interpolant
can always be built, as showed below:

\begin{proposition}\label{prop:positive-interpol}
  For any $u = a_1 \ldots a_n \in \Sigma^*$, if $\pi(u)$ is
  unsatisfiable then one can build an interpolant
  $\tuple{\top,I_0,\ldots,I_n,\bot}$ for $\pi(u)$, such that $I_i \in
  \Form^+(Q,\vec{x})$, for all $i \in [0,n]$.
\end{proposition}
\proof{ Let $\tuple{\top,I_0,\ldots,I_{n},\bot}$ be an interpolant,
  and $i \in [0,n]$ be the first index for which $I_i \not\in
  \Form^+(Q,\vec{x})$. By Proposition \ref{prop:asterisk}, we have
  that $\Delta^{i-1}(I_{i-1},a_i) \iffeq \exists Q_{i-1} ~.~
  I_{i-1}[Q_{i-1}/Q] \wedge \theta_i$. Since $I_{i-1} \in
  \Form^+(Q,\vec{x})$, by the choice of $i$, we have that
  $\Delta^{i-1}(I_{i-1},a_i) \in \Form^+(Q,\vec{x})$ as well. We
  replace $I_i$ with $\Delta^{i-1}(I_{i-1},a_i)$. The result is still
  an interpolant, because: \begin{compactitem}
  \item $I_{i-1}[Q_{i-1}/Q] \wedge \theta_i \models
    \Delta^{i-1}(I_{i-1},a_i)$, by Proposition \ref{prop:asterisk}, and
    %
  \item $\Delta^{i-1}(I_{i-1},a_i) \wedge \theta_{i+1} \models \exists
    Q_{i-1} ~.~ I_{i-1}[Q_{i-1}/Q] \wedge \theta_i \wedge \theta_{i+1}
    \models I_i[Q_i/Q] \wedge \theta_{i+1} \models I_{i+1}[Q_{i+1}/Q]$.
  \end{compactitem}
  We proceed with these replacements until there are no more formulae
  $I_i \not\in \Form^+(Q,\vec{x})$ left. \qed}

We recall that a run of $\A$ over $u$ is a sequence $\phi_0(Q),
\phi_1(Q,\vec{x}_0\cup\vec{x}_1), \ldots,
\phi_n(Q,\vec{x}_0\cup\ldots\cup\vec{x}_n)$, where $\phi_0$ is the
initial configuration $\iota$ and for each $k>0$, $\phi_k$ is obtained
from $\phi_{k-1}$ by replacing each state $q \in
\fv{\Bool}{\phi_{k-1}}$ by the formula
$\Delta(q,a_k)[\vec{x}_{k-1}/\overline{\vec{x}},\vec{x}_k/\vec{x}]$,
given by the transition function of $\A$. Observe that, because the
states are replaced with transition formulae when moving one step in a
run, the formulae thus produced lose track of the control history and
are not directly suitable to producing interpolants that relates
states and data.

The main idea behind the above definition of the interpolation problem
is that we would like to obtain an interpolant $\tuple{\top,I_0(Q),
  I_2(Q,\vec{x}), \ldots, I_{n}(Q,\vec{x}),\bot}$ whose formulae
\emph{combine states with the data constraints that must hold
  locally}, whenever the control reaches a certain boolean
configuration. This association of states with data valuations is
tantamount to defining efficient semi-algorithms, based on lazy
abstraction \cite{HJMS02}. Furthermore, the abstraction defined by the
interpolants generated in this way can also \emph{over-approximate the
  control structure} of an automaton, in addition to the sets of data
values encountered throughout its runs.

The correctness of this interpolation-based abstraction refinement
setup is captured by the progress property below, which guarantees
that adding the formulae of an interpolant for $\pi(u)$ to the set
$\preds$ of predicates suffices to exclude the spurious counterexample
$u$ from future searches.

\begin{lemma}\label{lemma:progress}
  For any non-empty sequence $u = a_1\ldots a_n \in \Sigma^*$
  such that $\Accept{\A}(u)$ is unsatisfiable, the following
  hold: \begin{compactenum}
  \item\label{it1:progress} $\Theta(u)$ is unsatisfiable, and
    %
  \item\label{it2:progress} if $\tuple{\top,I_0,\dots,I_{n},\bot}$ is
    an interpolant for $\Theta(u)$ such that $\set{I_i \mid i
      \in[0,n]} \subseteq \preds$ then $\AbsAccept{\A}(u)$ is
    unsatisfiable.
  \end{compactenum}
\end{lemma}
\proof{ Let $\Theta(u) \equiv \theta_0(Q_0) \wedge \theta_1(Q_0 \cup
  Q_1,\vec{x}_0 \cup \vec{x}_1) \wedge \ldots \wedge \theta_n(Q_{n-1}
  \cup Q_n,\vec{x}_{n-1} \cup \vec{x}_n) \wedge \theta_{n+1}(Q_n)$ in
  the following.

  (\ref{it1:progress}) We apply Proposition \ref{prop:asterisk}
  recursively and get: \[\Post{\A}(\iota,u)[Q_n/Q,\vec{x}_n/\vec{x}]
  \iff \exists Q_0 \ldots \exists Q_{n-1} \exists \vec{x}_0 \ldots
  \exists \vec{x}_{n-1} ~.~ \bigwedge_{i=0}^n \theta_i\] Assuming that
  $\Theta(u)$ is satisfiable, we obtain a model for $\Accept{\A}(u)
  \equiv \Post{\A}(\iota,u) \wedge \theta_{n+1}[Q/Q_n]$.

  (\ref{it2:progress}) If $\tuple{\top,I_0,\dots,I_n,\bot}$ is an
  interpolant for $\pi(u)$, the following entailments
  hold: \begin{compactitem}
  \item $\theta_0 \models I_0[Q_0/Q]$,
    %
  \item $I_{k-1}[Q_{k-1}/Q,\vec{x}_{k-1}/\vec{x}] \wedge \theta_k \models I_{k}[Q_k/Q,\vec{x}_k/\vec{x}]$, $\forall k\in[1,n]$. 
    %
  \item $I_{n}[Q_{n}/Q] \wedge \theta_{n+1} \models \bot$. 
  \end{compactitem}
  We prove that $\AbsPost{\A}(\iota,a_1\ldots a_n) \models I_{n}$ by
  induction on $n \geq 0$. This is sufficient to conclude because
  $\AbsAccept{\A}(a_1\ldots a_n) \equiv \AbsPost{\A}(\iota,a_1\ldots
  a_n) \wedge \theta_{n+1}[Q/Q_n] \models I_{n} \wedge
  \theta_{n+1}[Q/Q_n] \models \bot$. For the base case $n=0$, we have
  $\AbsPost{\A}(\iota,\varepsilon) \equiv \iota \equiv \theta_0[Q/Q_0]
  \models I_0$. For the induction step $n>0$, we compute:
  \[\begin{array}{l}
  \AbsPost{\A}(\iota,a_1\ldots a_n)[Q_n/Q] \equiv \text{ (by def. of $\AbsPost{\A}$)} \\
  \exists \vec{x}_{n-1} ~.~ \abs{\Delta^n(\AbsPost{\A}(\iota,a_1\ldots a_{n-1}),a_n)}[Q_n/Q] \models \text{ (by Prop. \ref{prop:asterisk})} \\
  \exists Q_{n-1}  \exists \vec{x}_{n-1} ~.~ \AbsPost{\A}(\iota,a_1\ldots a_{n-1})[Q_{n-1}/Q] \wedge \theta_n \models \text{ (ind. hyp.) } \\
  \exists Q_{n-1}  \exists \vec{x}_{n-1} ~.~ I_{n-1}[Q_{n-1}/Q] \wedge \theta_n \models I_n[Q_n/Q] \hfill\text{\qed}
  \end{array}\]}
We have now all the ingredients to describe the first emptiness
semi-algorithm for alternating data
automata. Algorithm\footnote{Though termination is not guaranteed, we
  call it algorithm for conciseness.} \ref{alg:predabs} builds an
\emph{abstract reachability tree} (ART) whose nodes are labeled with
formulae over-approximating the concrete sets of configurations, which
uses a covering relation between nodes in order to ensure that the set
of formulae labeling the nodes in the ART forms an antichain. Any
spurious counterexample is eliminated by computing an interpolant and
adding its formulae to the set of predicates.

\begin{algorithm}[t!]
{\scriptsize\begin{algorithmic}[0]
\State {\bf input}: an ADA $\A = \tuple{\vec{x},Q,\iota,F,\Delta}$
over the alphabet $\Sigma$ of input events

\State {\bf output}: $\true$ if $L(\A)=\emptyset$ and a data word $w
\in L(\A)$ otherwise
\end{algorithmic}}

{\scriptsize\begin{algorithmic}[1]

  \State let $\Art = \tuple{N,E,\rootNode,\Lambda,\lhd}$ be an ART

  \State initially $N = E = \lhd = \emptyset$, $\Lambda =
  \set{(\rootNode,\iota)}$, $\preds = \set{\bot}$, $\worklist =
  \tuple{\rootNode}$,

  \While {$\worklist \neq \emptyset$}
  \label{ln:while}

  \State dequeue $n$ from $\worklist$

  \State $N \leftarrow N \cup \set{n}$

  \If{$\Lambda(n) \wedge \bigwedge_{q \in Q \setminus F} (q
    \rightarrow \bot)$ is satisfiable} \Comment{counterexample candidate}

  \State let $u = a_1\ldots a_k$ be the label of the path from
  $\rootNode$ to $n$

  \If{$\Accept{\A}(u)$ is satisfiable} \Comment{counterexample is
    feasible}
  \label{ln:feasible}

  \State get model $\beta : Q \rightarrow \booli$, $\{\nu_i :
  \vec{x}_i \rightarrow \Data^\I \mid i=0\ldots k\}$ of $\Accept{\A}(u)$

  \State {\bf return} $w = (a_1,\nu_1) \ldots (a_k,\nu_k)$
  \label{ln:cex}
  \Comment{$w \in L(\A)$ by construction}

  \Else \Comment{spurious counterexample}

  \State $p \leftarrow \Call{FindPivot}{u,\Art}$
  \label{ln:pivot}

  \State let $v$ be the label of the path from $p$ to $n$ in $\Art$

  \State $\preds \leftarrow \preds \cup \set{I_0,\ldots,I_\ell}$,
  where $\tuple{\top,I_0,\ldots,I_\ell,\bot}$ is an interpolant for
  $\pivot{v}$

  \State let $\SubArt = \tuple{N',E',p,\Lambda',\lhd'}$ be the subtree
  of $\Art$ rooted at $p$

  \For{$(m,q) \in \lhd$ such that $q \in N'$}

  \State remove $m$ from $N$ and enqueue $m$ into $\worklist$

  \State remove $\SubArt$ from $\Art$

  \State enqueue $p$ into $\worklist$
  \label{ln:refine}
  \Comment{recompute the subtree rooted at $p$}

  \EndFor

  \EndIf

  \Else \label{ln:else}

  \For{$a \in \Sigma$}\label{ln:forall-events}
  \Comment{expand $n$}

  \State $\phi \leftarrow \AbsPost{\A}(\Lambda(n),a)$

  \If{exist $m \in N$ such that $\phi \models \Lambda(m)$} \label{ln:covered}

  \State $\lhd \leftarrow \lhd \cup \set{(n,m)}$ \Comment{$m$ covers $n$}
  \label{ln:direct-cover}

  \Else

  \State let $s$ be a fresh node

  \State $E \leftarrow E \cup \set{(n,a,s)}$

  \State $\Lambda \leftarrow \Lambda \cup \set{(s,\phi)}$

  \State $R \leftarrow \set{m \in \worklist \mid \Lambda(m) \models
    \phi}$ \Comment{worklist nodes covered by $s$}

  \For{$r \in R$}

  \For{$m \in N$ such that $(m,b,r) \in E$, $b \in \Sigma$}

  \State $\lhd \leftarrow \lhd \cup \set{(m,s)}$ 
  \label{ln:child-cover}
  \Comment{redirect covered children from $R$ into $s$}

  \EndFor

  \For{$(m,r) \in \lhd$}

  \State $\lhd \leftarrow \lhd \cup \set{(m,s)}$ 
  \label{ln:indirect-cover}
  \Comment{redirect covered nodes from $R$ into $s$}

  \EndFor

  \EndFor

  \State remove $R$ from $\Art$

  \State enqueue $s$ into $\worklist$ 
  \label{ln:expand}

  \EndIf

  \EndFor

  \EndIf

  \EndWhile

  \State {\bf return} $\true$
  \label{ln:true}
\end{algorithmic}}
\caption{Lazy Predicate Abstraction for ADA Emptiness}
\label{alg:predabs}
\end{algorithm}

Formally, an ART is tuple $\Art =
\tuple{N,E,\rootNode,\Lambda,R,T,\lhd}$, where: \begin{compactitem}
\item $N$ is a set of nodes, 
%
\item $E \subseteq N \times \Sigma \times
N$ is a set of edges, 
%
\item $\rootNode \in N$ is the root of the directed
tree $(N,E)$, 
%
\item $\Lambda : N \rightarrow \Form(Q,\vec{x})$ is a labeling of the
  nodes with formulae, such that $\Lambda(\rootNode) = \iota$,
%
\item $R : N \rightarrow 2^Q$ is a labeling of nodes with
\emph{replacement sets}, such that $R(\rootNode) = \fv{\Bool}{\iota}$,
%
\item $T : E \rightarrow \bigcup_{i=0}^\infty
\Form(Q_i,\vec{x}_i,Q_{i+1},\vec{x}_{i+1})$ is labeling of edges with
time-stamped formulae, and 
%
\item $\lhd \subseteq N \times N$ is a set of
\emph{covering edges}. 
\end{compactitem}
For simplicity, we write $T(n,a,m)$ for $T(e)$, when $e$ is $(n,a,m)$.
The replacement sets and time-stamped edge labels are defined below,
for each edge $e = (n,a,m) \in E$, if $n$ lies at depth\footnote{We
  consider that the root node has zero depth.} $k$ in the tree
$(N,E)$ then we have:
\[\begin{array}{rcl}
T(e) & = & \bigwedge_{q \in R(n)} (q_k \rightarrow \Delta(q,a)[Q_{k+1}/Q,\vec{x}_{k}/\overline{\vec{x}},\vec{x}_{k+1}/\vec{x}]) \\
R(m) & = & \bigcup_{q \in R(n)}\fv{\Bool}{\Delta(q,a)}
\end{array}\]
Each node $n \in N$ corresponds to a unique path from the root to $n$,
labeled by a sequence $\lambda(n) \in \Sigma^*$ of input events. The
least unfeasible suffix of $\lambda(n)$ is the smallest sequence $v =
a_1 \ldots a_k$, such that $\lambda(n) = wv$, for some $w \in
\Sigma^*$ and the formula (\ref{eq:pivot}) below is unsatisfiable. The
\emph{pivot} of $n$ is the node $p$ corresponding to the start of $v$
in $\lambda(n)$.
\begin{equation}\label{eq:pivot}
  \pivot{v} \equiv \Lambda(p)[Q_0/Q] \wedge
  \theta_1(Q_0 \cup Q_1,\vec{x}_0 \cup \vec{x}_1) \wedge \ldots \wedge
  \theta_{k+1}(Q_k)
\end{equation}
where $\theta_1,\ldots,\theta_{k+1}$ are defined as in
(\ref{eq:interpolation-problem}), for $\theta_0 \equiv
\Lambda(p)[Q_0/Q]$.

With these considerations, Algorithm \ref{alg:predabs} uses a worklist
iteration to build an ART. We keep newly expanded nodes of $\Art$ in a
queue $\worklist$, thus implementing a breadth-first exploration
strategy, which guarantees that the shortest counterexamples are
explored first. When the search encounters a counterexample candidate
$u$, it is checked for spuriousness. If the counterexample is
feasible, the procedure returns a data word $w \in L(\A)$, which
interleaves the input events of $u$ with the data valuations from the
model of $\Accept{\A}(u)$ (since $u$ is feasible, clearly
$\Accept{\A}(u)$ is satisfiable). Otherwise, if $u$ is spurious, we
compute its pivot $p$ (line \ref{ln:pivot}), add the interpolants for
the least unfeasible suffix of $u$ to the set of predicates $\preds$,
remove and recompute the subtree of $\Art$ rooted at $p$.

Termination of Algorithm \ref{alg:predabs} depends on the ability of a
given interpolating decision procedure for the combined boolean and
data theory $\theory(\sig,\I)$ to provide interpolants that yield a
safety invariant, whenever $L(\A) = \emptyset$. In this case, we
use the covering relation $\lhd$ to ensure that, when a newly
generated node is covered by a node already in $N$, it is not added to
the worklist, thus cutting the current branch of the search. Formally,
for any two nodes $n,m \in N$, we have $n \lhd m$ iff
$\AbsPost{\A}(\Lambda(n),a) \models \Lambda(m)$ for some $a \in
\Sigma$, in other words, if $n$ has a successor whose label entails
the label of $m$. The correctness criteria of Algorithm
\ref{alg:predabs} are showed in the following theorem:

\begin{theorem}\label{thm:predabs}
  Given an automaton $\A$, such that $L(\A) \neq \emptyset$, Algorithm
  \ref{alg:predabs} terminates and returns a word $w \in L(\A)$. If
  Algorithm \ref{alg:predabs} terminates reporting $\true$, then
  $L(\A) = \emptyset$.
\end{theorem}
\proof{We prove the following invariant: each time Algorithm
  \ref{alg:predabs} reaches line \ref{ln:while}, the set $W$ of nodes
  in $\worklist$ contains all the frontier nodes in the ART
  $\tuple{N\cup W,E,\rootNode,\Lambda,\lhd}$ which are not covered by
  some node in $N$, namely that:
  \begin{equation}\label{inv:predabs}
  W = \set{n \not\in N \mid \forall m \in N~ \forall a \in \Sigma ~.~ (n,a,m)
    \not\in E ~\wedge~ (n,m) \not\in \lhd}
  \end{equation}
  Initially, this is the case because $W = \set{\rootNode}$ and $N = E
  = \lhd = \emptyset$. If the invariant holds previously, at line
  \ref{ln:while}, it will hold again after line \ref{ln:refine} is
  executed, because, when the subtree rooted at the pivot $p$ is
  removed, $p$ becomes a member of the set of uncovered frontier
  nodes, and is added to $W$ at line \ref{ln:refine}. The last case
  considers that the invariant holds at line \ref{ln:while} and the
  control follows the else branch at line \ref{ln:else}. In this case,
  the newly created frontier node $s$ is added to $W$ only if it is
  not covered by an existing node in $N$ (line \ref{ln:covered}).

  Next we prove that, if Algorithm \ref{alg:predabs} returns $\true$,
  then $\bigvee_{n \in N} \Lambda(n)$ defines a safety
  invariant. Suppose that Algorithm \ref{alg:predabs} returns at line
  \ref{ln:true}. Then it must be that $W = \emptyset$. Because
  (\ref{inv:predabs}) is invariant, each node in $N$ is either covered
  by another node in $N$, or all its successors are in $N$. We prove
  first that $\bigvee_{n \in N} \Lambda(n)$ is an invariant: for any
  $u \in \Sigma^*$, there exists some node $n \in N$ such that
  $\Post{\A}(\iota,u) \models \Lambda(n)$. Let $u \in \Sigma^*$ be an
  arbitrary sequence. If $u$ labels the path from $\rootNode$ to some
  $n \in N$, we have $\Post{\A}(\iota,u) \models \AbsPost{\A}(\iota,u)
  \models \Lambda(n)$ and we are done. Otherwise, let $v$ be the
  (possibly empty) prefix of $u$ which labels the path from
  $\rootNode$ to some $n \in N$, which is covered by another $m \in
  N$, where $(n,a,m) \in E$, that is $u = vav'$, for some $a \in
  \Sigma$ and $v'\in\Sigma^*$. Moreover, we have $\Post{\A}(\iota,va)
  \models \AbsPost{\A}(\iota,va) \models \Lambda(m)$, by the
  construction of the set $\lhd$ of covering edges --- lines
  \ref{ln:direct-cover}, \ref{ln:child-cover},
  \ref{ln:indirect-cover}. Continuing this argument recursively from
  $m$, since $\len{v'} < \len{u}$, we shall eventually discover a node
  $p$ such that $\Post{\A}(\iota,u) \models \Lambda(p)$.

  To prove that $\bigvee_{n \in N} \Lambda(n)$ is, moreover, a safety
  invariant, suppose, by contradiction, that there exists $u \in
  \Sigma^*$ such that $\Accept{\A}(u)$ is satisfiable. By the previous
  point, there exists a node $p \in N$ such that $\Post{\A}(\iota,u)
  \models \Lambda(p)$. But then we have $\Accept{\A}(\iota,u) \models
  \Accept{\A}(\iota,\lambda(p))$, thus $\Accept{\A}(\iota,\lambda(p))$
  is satisfiable as well. However, this cannot be the case, because
  $p$ has been processed at line \ref{ln:feasible} and Algorithm
  \ref{alg:predabs} would have returned a counterexample,
  contradicting the assumption that it returns $\true$. This concludes
  the proof that $\bigvee_{n\in N} \Lambda(n)$ is a safety invariant,
  thus $L(\A) = \emptyset$, by Lemma \ref{lemma:safety-invariant}.

  We have then proved the second point of the statement. For the first
  point, assume that $L(\A) \neq 0$ and let $w = (a_1,\nu_1) \ldots
  (a_k,\nu_k) \in L(\A)$ be a word. By the above, Algorithm
  \ref{alg:predabs} cannot return $\true$. Suppose, by contradiction
  that it does not terminate. Since the sequences from $\Sigma^*$ are
  explored in breadth-first order, every sequence of length $k$ is
  eventually explored, which leads to the discovery of $w$ at line
  \ref{ln:feasible}. Then Algorithm \ref{alg:predabs} terminates
  returning $w \in L(\A)$. \qed}

\section{Checking ADA Emptiness with \impact}

As pointed out by a number of authors, the bottleneck of predicate
abstraction is the high cost of reconstructing parts of the ART,
subsequent to the refinement of the set of predicates. The main idea
of the \impact~ procedure \cite{mcmillan06} is that this can be
avoided and the refinement (strenghtening of the node labels of the
ART) can be performed in-place. This refinement step requires an
update of the covering relation, because a node that used to cover
another node might not cover it after the strenghtening of its label.

%% The ART built by Algorithm \ref{alg:impact} corresponds to an
%% unfolding of the input automaton by enumerating the input event
%% sequences in $\Sigma^*$. Each time a new node is created, its label is
%% initially $\top$. This label is subject to subsequent strenghtening,
%% when a spurious counterexample is encountered. To speedup the
%% computation of interpolants for spurious counterexamples, we use two
%% additional labeling functions for bookkeeping. Formally, the ART
%% structure used by Algorithm \ref{alg:impact} is defined as $\Art =
%% \tuple{N,E,\rootNode,\Lambda,R,T,\lhd}$, where $N,E,\Lambda$ and
%% $\lhd$ are defined as for Algorithm \ref{alg:predabs} and,
%% moreover: \begin{compactitem}
%% \item $R : N \rightarrow 2^Q$ is a labeling of nodes with replacement
%%   sets, such that $R(\rootNode) = \fv{\Bool}{\iota}$, 
%% %
%% \item $T : E \rightarrow \bigcup_{i=0}^\infty
%%   \Form(Q_i,\vec{x}_i,Q_{i+1},\vec{x}_{i+1})$ is labeling of edges
%%   with time-stamped formulae, and
%% %
%% \item for each edge $e = (n,a,m) \in E$, if $n$ lies at
%%   depth\footnote{We consider that the root node has zero depth.}  $k$
%%   in the tree $(N,E)$, then we have:
%% \end{compactitem}
%%   \[\begin{array}{rcl}
%%   T(e) & = & \Theta(n,a,k) \\
%%   R(m) & = & \bigcup_{q \in R(n)}\fv{\Bool}{\Delta(q,a)} \text{, where} \\
%%   \Theta(n,a,k) & \equiv & \bigwedge_{q \in R(n)} (q_k \rightarrow
%%   \Delta(q,a)[Q_{k+1}/Q,\vec{x}_{k}/\overline{\vec{x}},\vec{x}_{k+1}/\vec{x}])
%%   \end{array}\]
%% For simplicity, we write $T(n,a,m)$ for $T(e)$, when $e$ is
%% $(n,a,m)$. This additional bookkeeping is used as follows. Given a
%% sequence $u = a_1\ldots a_k \in \Sigma^*$ that labels the path from
%% the root to a node $n$ in $\Art$, the conjunction:
%% \[\mu(n) \equiv \Lambda(\rootNode) \wedge T(\rootNode,a_1,n_1) \wedge \ldots
%% T(n_{k-1},a_k,n_k) \wedge \bigwedge_{q \in R(n_k) \setminus F}(q_k
%% \rightarrow \bot)\] is equivalent to the interpolation problem
%% $\pi(u)$ and can be directly computed by a traversal of the path from
%% $\rootNode$ to $n$.

We consider a total alphabetical order $\prec$ on $\Sigma$ and lift it
to the total lexicographical order $\prec^*$ on $\Sigma^*$. We write
$n \prec^* m$ if the sequence labeling the path from $\rootNode$ to
$n$ is strictly smaller, in the lexicographical order, than the
sequence labeling the path from $\rootNode$ to $m$.

A node $n \in N$ is \emph{covered} if $(n,p) \in \lhd$ or it has an
ancestor $m$ such that $(m,p) \in \lhd$, for some $p \in N$. A node
$n$ is \emph{closed} if it is covered, or $\Lambda(n) \not\models
\Lambda(m)$ for all $m \in N$ such that $n \prec^* m$.

\begin{algorithm}[t!]
{\scriptsize\begin{algorithmic}[0]
  \State {\bf input}: an ADA $\A = \tuple{\vec{x},Q,\iota,F,\Delta}$
  over the alphabet $\Sigma$ of input events

  \State {\bf output}: $\true$ if $L(\A)=\emptyset$ and a data word $w
  \in L(\A)$ otherwise
\end{algorithmic}}

{\scriptsize\begin{algorithmic}[1]

  \State let $\Art = \tuple{N,E,\rootNode,\Lambda,R,T,\lhd}$ be an ART

  \State initially $N = E = T = \lhd = \emptyset$, $\Lambda =
  \set{(\rootNode,\iota)}$, $R = \fv{\Bool}{\iota[Q_0/Q]}$,
  $\worklist=\set{\rootNode}$

  \While{$\worklist \neq \emptyset$}
  \label{ln:impact-while}

  \State dequeue $n$ from $\worklist$ 
  \label{ln:impact-dequeue}

  \State $N \leftarrow N \cup \set{n}$

  \State let $(\rootNode,a_1,n_1),(n_1,a_2,n_2), \ldots,
  (n_{k-1},a_k,n)$ be the path from $\rootNode$ to $n$

  \If{$\Phi(a_1 \ldots a_k)$ is satisfiable} 
  \Comment{counterexample is feasible}

  \State get model $\beta : Q_i \rightarrow \booli$, $\{\nu_i :
  \vec{x}_i \rightarrow \Data^\I \mid i=0\ldots k\}$ of $\mu(n)$

  \State {\bf return} $w = (a_1,\nu_1) \ldots (a_k,\nu_k)$
  \label{ln:impact-nonempty}
  \Comment{$w \in L(\A)$ by construction}

  \Else \Comment{spurious counterexample}

  \State let $\tuple{\top,I_0,\ldots,I_k,\bot}$ be an interpolant for
  $\Phi(a_1 \ldots a_k)$
  \label{ln:refine-begin}

  \State let $b$ be $\false$

  \For{$i=0,\ldots,k$}

  \If{$\Lambda(n_i) \not\models I_i$}

  \State $\lhd \leftarrow \lhd \setminus \set{(m,n_i) \in \lhd \mid m \in N}$

  \State $\Lambda(n_i) \leftarrow \Lambda(n_i) \wedge I_i$
  \Comment{strenghten the label of $n_i$}

  \If{$\neg b$}

  \State $b \leftarrow \Call{Close}{n_i}$
  \label{ln:refine-end}

  \EndIf % $\neg b$

  \EndIf % $\Lambda(n_i) \not\models I_i$

  \EndFor

  \EndIf % $\mu(n)$ is satisfiable

  \If{$n$ is not covered}

  \For{$a \in \Sigma$} 
  \Comment{expand $n$}
  \label{ln:expand-begin}

  \State let $s$ be a fresh node and $e = (n,a,s)$ be a new edge

  \State $E \leftarrow E \cup \set{e}$
  \label{ln:edge-insert}

  \State $\Lambda \leftarrow \Lambda \cup \set{(s,\top)}$

  \State $T \leftarrow T \cup \set{(e, \Theta(n,a,k))}$

  \State $R \leftarrow R \cup \{(s,\bigcup_{q \in R(n)}\fv{\Bool}{\Delta(q,a)})\}$

  \State enqueue $s$ into $\worklist$
  \label{ln:expand-end}

  \EndFor

  \EndIf % $n$ is not covered

  \EndWhile

  \State {\bf return} $\true$
\end{algorithmic}

\begin{algorithmic}[1]
\Function{Close}{$x$} {\bf returns} $\Bool$

\For{$y \in N$ such that $y \prec^* x$}

\If{$\Lambda(x) \models \Lambda(y)$}

\State $\lhd \leftarrow \left(\lhd \setminus \set{ (p,q) \in \lhd \mid
  \text{$q$ is $x$ or a successor of $x$}}\right) \cup \set{(x,y)}$
\label{ln:close-uncover}

\State {\bf return} $\true$

\EndIf

\EndFor

\State {\bf return} $\false$

\EndFunction
\end{algorithmic}}
\caption{\impact~ for ADA Emptiness}
\label{alg:impact}
\end{algorithm}

The execution of Algorithm \ref{alg:impact} consists of three
phases\footnote{These phases correspond to the \textsc{Close},
  \textsc{Refine} and \textsc{Expand} procedures from the original
  description of \impact~\cite{mcmillan06}.}: \emph{close},
\emph{refine} and \emph{expand}. Let $n$ be a node removed from the
worklist at line \ref{ln:impact-dequeue}. If $\Phi(\lambda(n))$ is
satisfiable, the counterexample $\lambda(n)$ is feasible, in which
case a model of $\Phi(\lambda(n))$ is obtained and a word $w \in
L(\A)$ is returned. 
%
%% When a node $n$ is removed from the worklist, it is first closed (line
%% \ref{ln:close-worklist}) by finding a lexicographically bigger node
%% $m$ such that $\Lambda(n) \models \Lambda(m)$, if one exists, and
%% adding the pair $(n,m)$ to $\lhd$. As in Algorithm \ref{alg:predabs},
%% the intuition is that for any feasible counterexample that can be
%% discovered from $n$ there exists a feasible counterexample stemming in
%% $m$, thus we do not need to process $n$ further.
%
%% If $n$ is not covered, we check if it is the endpoint of a
%% counterexample candidate, which can only be the case when the
%% replacement set $R(n)$ contains at least one final state (line
%% \ref{ln:cex}). Observe that, if $u$ is the label of the path from
%% $\rootNode$ to $n$, then $\Accept{\A}(u)$ is satisfiable only if it
%% contains a (positive occurrence of) a final state. The counterexample
%% is moreover feasible iff $\mu(n)$ is satisfiable, in which case a
%% model of $\mu(n)$ is obtained and a word $w \in L(\A)$ returned.
%
Otherwise, $\lambda(n)$ is a spurious counterexample and the procedure
enters the refinement phase (lines
\ref{ln:refine-begin}-\ref{ln:refine-end}). The interpolant for
$\Phi(\lambda(n))$ is used to strenghten the labels of all the
ancestors of $n$, by conjoining the formulae of the interpolant to the
existing labels.

In this process, the nodes on the path between $\rootNode$ and $n$,
including $n$, might become eligible for coverage, therefore we
attempt to close each ancestor of $n$ that is impacted by the
refinement (line \ref{ln:refine-end}). Observe that, in this case the
call to $\Call{Close}{}$ must uncover each node which is covered by a
successor of $n$ (line \ref{ln:close-uncover} in the $\Call{Close}{}$
procedure). This is required because, due to the over-approximation of
the sets of reachable configurations, the covering relation is not
transitive, as explained in \cite{mcmillan06}. If $\Call{Close}{}$
adds a covering edge $(n_i,m)$ to $\lhd$, it does not have to be
called for the successors of $n_i$ on this path, which is handled via
the boolean flag $b$.

Finally, if $n$ is still uncovered (it has not been previously covered
during the refinement phase) we expand $n$ (lines
\ref{ln:expand-begin}-\ref{ln:expand-end}) by creating a new node for
each successor $s$ via the event $a \in \Sigma$ and inserting it into
the worklist.

The following lemma states that the labeling of the nodes in an ART
built by Algorithm \ref{alg:impact} is consistent with the semantics
of the transition function of $\A$.

\begin{lemma}\label{lemma:well-labeled}
  Given an ART $\Art = \tuple{N,E,\rootNode,\Lambda,R,T,\lhd}$ built
  by Algorithm \ref{alg:impact}, $\Post{\A}(\Lambda(n),a) \models
  \Lambda(m)$, for all $(n,a,m) \in E$.
\end{lemma}
\proof{ We distinguish two cases. First, if $(n,a,m)$ occurs on a path
  in $\Art$ that has never been refined, then $\Lambda(m)=\top$ and
  the entailment holds trivially. Otherwise, let $\Omega$ be the set
  of paths $\omega = (n_0,a_1,n_1), \ldots, (n_{k-1},a_k,n_k)$, where
  $n_0 = \rootNode$ and $(n,a,m) = (n_{i-1},a_i,n_i)$, for some $i \in
  [1,k]$ and, moreover, $a_1\ldots a_k$ was found, at some point, to
  be a spurious counterexample. Let
  $\langle\top,I^\omega_0,\ldots,I^\omega_k\equiv\bot\rangle$ be an
  interpolant for $\Phi(a_1 \ldots a_k) \equiv \Lambda(\rootNode)
  \wedge \bigwedge_{i=1}^k \theta_i \wedge \bigwedge_{q \in R(n_k)}
  (q_k \rightarrow \bot)$, such that $I^\omega_i \in
  \Form^+(Q,\vec{x})$, for all $i \in [0,k]$. According to Proposition
  \ref{prop:positive-interpol}, it is possible to build such an
  interpolant, when $\Phi(a_1 \ldots a_k)$ is unsatisfiable. By
  Proposition \ref{prop:asterisk}, we obtain
  $\Delta^i(I^\omega_{i-1},a_i)[Q_i/Q] \iffeq \exists Q_{i-1} ~.~
  I^\omega_{i-1}[Q_{i-1}/Q,\vec{x}_{i-1}/\vec{x}] \wedge \theta_i$
  and, since $I^\omega_{i-1}[Q_{i-1}/Q,\vec{x}_{i-1}/\vec{x}] \wedge
  \theta_i \models I^\omega_i[Q_i/Q,\vec{x}_i/\vec{x}]$, we obtain
  that $\Delta^i(I^\omega_{i-1},a_i)[Q_i/Q] \models
  I^\omega_i[Q_i/Q,\vec{x}_i/\vec{x}]$. Since $\Lambda(n_{i-1}) =
  \bigwedge_{\omega \in \Omega} I^\omega_{i-1}$ and $\Lambda(n_{i}) =
  \bigwedge_{\omega \in \Omega} I^\omega_{i}$, we obtain
  $\Post{\A}(\Lambda(n_{i-1}),a_i) \models \Lambda(n_i)$. \qed}

The correctness of Algorithm \ref{alg:impact} is proved in the
following.

\begin{theorem}\label{thm:impact}
  Given an automaton $\A$, such that $L(\A) \neq \emptyset$, Algorithm
  \ref{alg:impact} terminates and returns a word $w \in L(\A)$. If
  Algorithm \ref{alg:impact} terminates reporting $\true$, then
  $L(\A) = \emptyset$.
\end{theorem}
\proof{ We prove first that, each time Algorithm \ref{alg:impact}
  reaches the line \ref{ln:impact-while}, we have:
  \begin{equation}\label{inv:impact}
  W = \set{n \not\in N \mid n \text{ uncovered, } \exists a \in
    \Sigma~ \forall s \in N ~.~ (n,a,s) \not\in E}
  \end{equation}
  Initially, $W = \set{\rootNode}$ and $N=E=\lhd=\emptyset$, thus
  (\ref{inv:impact}) holds trivially. Suppose that (\ref{inv:impact})
  holds at when reaching line \ref{ln:impact-while} and some node $n$
  was removed from $W$ and inserted into $N$. We distinguish two
  cases, either: \begin{itemize}
  \item $n$ is covered, in which case $W$ becomes $W \setminus
    \set{n}$, or
  %
  \item $n$ is not covered, in which case $W$ becomes $(W \setminus
    \set{n}) \cup S$, where $S = \set{s \not\in N \mid (n,a,s) \in E,
      a \in \Sigma}$ is the set of fresh successors of $n$. But then
    no node $s \in S$ is covered and has successors in $E$.
  \end{itemize}
  Then the condition (\ref{inv:impact}) holds next time line
  \ref{ln:impact-while} is reached, thus it is invariant.

  Suppose first that Algorithm \ref{alg:impact} returns $\true$, thus
  $W = \emptyset$ and, by (\ref{inv:impact}), for each node in $n \in
  N$ one of the following hold: \begin{compactitem}
  \item $n$ is covered, or
  \item for each $a \in \Sigma$ there exists $s \in N$ such that
    $(n,a,s) \in E$. 
  \end{compactitem}
  We prove that, in this case, $\bigvee_{n \in N} \Lambda(n)$ defines
  a safety invariant and conclude that $L(\A) = \emptyset$, by Lemma
  \ref{lemma:safety-invariant}. To this end, let $u = a_1 \ldots a_k
  \in \Sigma^*$ be an arbitrary sequence and let $v_1$ be the largest
  prefix of $u$ that labels a path from $\rootNode$ to some node $n_1
  \in N$. If $v_1 = u$ we are done. Otherwise, by the choice of $v_1$,
  it must be the case that a successor of $n_1$ is missing from
  $(N,E)$, thus $n_1$ must be covered. Let $n'_1$ be the closest
  ancestor of $n_1$ such that $(n'_1,n''_1) \in \lhd$, for some $n''_1
  \in N$, and let $v'_1$ be the prefix of $v_1$ leading to $n'_1$. By
  the construction of $\lhd$ (line \ref{ln:close-uncover} in function
  $\Call{Close}{}$), we have $\Lambda(n'_1) \models
  \Lambda(n''_1)$. Applying Lemma \ref{lemma:well-labeled} inductively
  on $v'_1$, we obtain that $\Post{\A}(\iota,v'_1) \models
  \Lambda(n'_1)$, thus $\Post{\A}(\iota,v'_1) \models \Lambda(n''_1)$.
  Continuing inductively from $n''_1$, we exhibit a sequence of
  strings $v'_1,\ldots,v'_\ell \in \Sigma^*$ and nodes $\rootNode =
  m_0,m_1,\ldots,m_\ell$ such that, for all $i \in
  [1,\ell]$: \begin{compactitem}
  \item $v'_i$ labels the path between $m_{i-1}$ and $m_i$ in $(N,E)$,    
    %
  \item $\Post{\A}(\iota,v'_1\ldots v'_i) \models \Lambda(m_i)$.
  \end{compactitem}
  Moreover, we have $u = v'_1 \ldots v'_\ell$, thus
  $\Post{\A}(\iota,u) \models \Lambda(m_\ell)$ and we are done showing
  that $\bigvee_{n\in N} \Lambda(n)$ is an invariant.

  To prove that $\bigvee_{n \in N} \Lambda(n)$ is, moreover, a safety
  invariant, suppose that $\Accept{\A}(u)$ is satisfiable, for some $u
  \in \Sigma^*$ and let $n \in N$ be a node such that
  $\Post{\A}(\iota,u) \models \Lambda(n)$. By the previous point, such
  a node must exist. But then $\Accept{\A}(u) \models
  \Accept{\A}(\lambda(n))$, thus $\Accept{\A}(\lambda(n))$ is
  satisfiable, and Algorithm \ref{alg:impact} returns at line
  \ref{ln:impact-nonempty}, upon encountering $\lambda(n)$. But this
  contradicts the assumption that Algorithm \ref{alg:impact} returns
  $\true$, hence we have proved that $\bigvee_{n \in N} \Lambda(n)$ is
  a safety invariant, and $L(\A) = \emptyset$ follows, by Lemma
  \ref{lemma:safety-invariant}. We have then proved the second point
  of the statement.

  To prove the first point, assume that $L(\A) \neq \emptyset$. By the
  previous point, Algorithm \ref{alg:impact} does not return
  $\true$. Suppose, by contradiction, that it does not terminate and
  conclude using the breadth-first exploration argument from the proof
  of Theorem \ref{thm:predabs}. \qed}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\bibliographystyle{abbrv} \bibliography{refs}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\end{document}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%