### 2017-05-16 predicate-automata.tex
ok

\documentclass{llncs}

\usepackage[usenames,dvipsnames]{color}

\usepackage{latexsym}
\usepackage{amsxtra} 
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{pslatex}
% \usepackage{epsfig}
\usepackage{wrapfig}
\usepackage{paralist}
\usepackage{stmaryrd}
\usepackage{txfonts}
\usepackage{framed}
\usepackage{makecell}
\usepackage{url}
\usepackage{tikz}
\usetikzlibrary{automata,positioning, calc}
\usepackage[inline,shortlabels]{enumitem}

\usepackage{proof}

\usepackage{algorithm}
\usepackage{algorithmicx}
\usepackage[noend]{algpseudocode}

\usepackage[draft]{commenting}

\pagestyle{plain}

\include{commands}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{document}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\title{From Predicate Automata to Alternating Data Automata}

\author{Radu Iosif}
\institute{CNRS, Verimag \\
Radu.Iosif@univ-grenoble-alpes.fr}

\maketitle

\section{Definitions}

Let $Q$ be a finite set of predicates, where for each $q \in Q$, we
denote by $\#(q)\geq0$ its arity. We denote by $Q_0$ the subset of $Q$
consisting of predicates of arity zero. By $\Sigma$ we denote a
finite set of symbols, called alphabet.

We consider an infinite countable set $\vars$ of first-order
variables, ranging over $\nat$, and we associate each predicate $q \in
Q$ a tuple of arguments $\vec{x}_q = \tuple{x_{q,1}, \ldots,
  x_{q,\#(q)}}$ such that, for any two $p,q \in Q$, $\vec{x}_p$ and
$\vec{x}_q$ have no elements in common, when $p \neq q$.

\section{Translation of PA into ADA}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\end{document}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%