### 2017-05-17 predicate-automata.tex
ok

\documentclass{llncs}

\usepackage[usenames,dvipsnames]{color}

\usepackage{latexsym}
\usepackage{amsxtra} 
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{pslatex}
% \usepackage{epsfig}
\usepackage{wrapfig}
\usepackage{paralist}
\usepackage{stmaryrd}
\usepackage{txfonts}
\usepackage{framed}
\usepackage{makecell}
\usepackage{url}
\usepackage{tikz}
\usetikzlibrary{automata,positioning, calc}
\usepackage[inline,shortlabels]{enumitem}

\usepackage{proof}

\usepackage{algorithm}
\usepackage{algorithmicx}
\usepackage[noend]{algpseudocode}

\usepackage[draft]{commenting}

\pagestyle{plain}

\include{commands-lncs}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{document}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\title{From Predicate Automata to Alternating Data Automata}

\author{Radu Iosif}
\institute{CNRS, Verimag \\
Radu.Iosif@univ-grenoble-alpes.fr}

\maketitle

\section{Definitions}

Let $Q$ be a finite set of predicates, where for each $q \in Q$, we
denote by $\#(q)\geq0$ its arity. We denote by $Q_0$ the subset of $Q$
consisting of predicates of arity zero. By $\Sigma$ we denote a finite
set of symbols, disjoint from $Q$, called an alphabet.

We consider an infinite countable set $\vars$ of first-order
variables, ranging over $\nat$. For a tuple of variables $\vec{x} =
\tuple{x_1,\ldots,x_n}$, we define a one-to-one corresponding tuple
$\overline{\vec{x}} = \tuple{\overline{x_1},\ldots,\vec{x_n}}$. We
associate each predicate $q \in Q$ a tuple of arguments $\vec{x}_q =
\tuple{x_{q,1}, \ldots, x_{q,\#(q)}}$ such that, for any two $p,q \in
Q$, if $p \neq q$ then any two of $\vec{x}_p$, $\overline{\vec{x}}_p$,
$\vec{x}_q$ and $\overline{\vec{x}}_q$ have no elements in common,
respectively.

An \emph{atom} is a predicate $q(\vec{x}_q)$,
$q(\overline{\vec{x}}_q)$, $x \teq y$ or $x \not\teq y$, for $x,y \in
\vars$ and a \emph{formula} is a positive boolean combination of
atoms. We denote by $\Form(Q,\vec{x})$ the set of formulae with
predicates from $Q$ and variables from $\vec{x}$.

\begin{definition}[Predicate Automata]
\end{definition}

\section{Translation of PA into ADA}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\end{document}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%