### 2017-09-12 rebuttal.txt
ok

We are grateful to all reviewers for their great work. The main points
of criticism are addressed as follows:

1. Why are our semi-algorithms performing slower than Includer [11] ?

First of all, Includer is specifically designed to solve language
inclusion problems between nondeterministic data automata, which are a
(significant) special case of alternation, whereas the semi-algorithms
presented in this paper can deal with the entire class of alternating
data automata, not just the ones resulting from inclusion problems. It
is common that procedures for more general problems are slower than
the ones for specific problems, although ideally, it is desirable to
have matching speed on the same class of problems.

As reviewer 2 points out, the generalized subset construction used by
Includer does not cause a huge blowup for the examples considered in
this paper and in [11]. Moreover, Includer uses the LIA interpolation
procedure of MathSAT5, which seems to be faster than the more generic
UFLIA interpolation of Z3 used here (UFLIA is needed in order to deal
with the boolean variables representing the states). In some cases, Z3
answers unsat almost immediately and then takes up to 5 seconds to
extract the interpolant. We believe that further experimenting with
different Z3 tactics and tuning the encoding of the interpolation
procedure, as well as considering a larger set of benchmarks than the
preliminary benchmarks used to test our first prototype will greatly
improve the performance of our tool.

2. What is the difference between the semi-algorithms in this paper
and the original ones of Blast [8] and Impact [15] ?

Both Blast and Impact are designed for checking emptiness (safety) of
nondeterministic data automata (programs), whereas our methods address
the emptiness problem for alternating data automata. A first
difference between our methods and [8,15] is a more sophisticated
Abstract Reachability Tree (ART) data structure (page 5, 2nd paragraph
of section 5), which requires labeling both nodes and edges with
formulae. However, on the positive side, our encoding of path formulae
(interpolation problems) allows the basic principles of [8,15] to be
carried over to the alternating setting (we represent path formulae as
conjunctions of implications and abstract sets of configurations as
combinations of boolean and data theory atoms), with some specific
differences listed below:

2.a) Difference with Blast [8]. Algorithm 1 (page 6) uses antichains
to prune the search space (lines 29-35). Using antichains is a
specific technique for classical finite-alphabet alternating
automata. We adapt this idea from [5] to optimize the predicate
abstraction emptiness checking of ADA.

2.b) Difference with Impact [15]. The original description of Impact
[15] is based on a nondeterministic application of the Cover, Refine
and Expand procedures, whereas Algorithm 2 (page 8) is a detailed
deterministic procedure that closely mirrors our implementation. In
particular, we chose to cover nodes lazily, only if their labels are
strenghtened by refinement (line 18).

3. Using the Lyndon interpolation theorem instead of post-images to
prove Proposition 1. This is a very interesting remark that we did not
consider initially and which allows to use positive interpolants
directly instead of post-images (quantifiers can be eliminated if the
data theory is LIA, but quantifier elimination has a non-negligeable
cost). If accepted, we are willing to change this point in the final
version, provided that we can find a published reference for the fact
that McMillan's Z3 interpolation procedure fits with Lyndon's theorem
(their CAV13, FMCAD11 or TCS05 papers on interpolation do not seem to
refer to it).

4. Several reviewers expressed concerns relative to the paper being
too theoretical and out of the scope of FMCAD. As authors of a paper
on formal verification techniques, we felt the need to define and
prove each point formally. However, the paper does not address issues
such as decidability or computational complexity of subclasses of ADA,
which could fit a more theoretical conference. Although interpolation
can be used as an argument for decidability (when suitable
restrictions are imposed, such as for symbolic alternating automata)
it is perhaps not a very good framework to derive complexity
bounds. In general, theoretical complexity proofs require different
techniques (e.g. reductions to other decidable problems whose
complexity is known) which are clearly out of the scope of the present
paper.