### 2019-02-07 quantifiers/appendix.tex

\section{Emptiness Semi-algorithm for First Order Alternating Automata}
\label{app:algorithm}

\section{Proofs}
\label{app:proofs}