### 2019-02-07 quantifiers/appendix.tex
ok

\section{Emptiness Checking for First Order Alternating Automata}
\label{app:algorithm}

\section{Proofs of the Results in the Paper}
\label{app:proofs}