### 2019-02-07 quantifiers/appendix.tex
ok

\section{Emptiness Checking for First Order Alternating Automata}
\label{app:algorithm}

The execution of Algorithm \ref{alg:impact} consists of three phases,
corresponding to the \textsc{Close}, \textsc{Refine} and
\textsc{Expand} of the original \impact\ procedure \cite{McMillan06}.

\begin{algorithm}[htb]
\begin{algorithmic}[0] \State {\bf input}: a first order
  alternating automaton $\A = \tuple{\Sigma,X,Q,\iota,F,\Delta}$

  \State {\bf output}: $\top$ if $L(\A)=\emptyset$, or word $w \in
  L(\A)$, otherwise

  \State {\bf data structures}: $\worklist$ and unfolding tree
  $\mathcal{U} = \tuple{N,E,\rootNode,U,\lhd}$,
  where: \begin{compactitem}
  \item $N$ is a set of nodes,
  \item $E \subseteq N \times \Sigma \times N$ is a set of edges labeled by input events,
  \item $U : N \rightarrow \posforms(Q,\emptyset)$ is a labeling of
    nodes with positive sentences
  \item $\lhd \subseteq N \times N$ is a coverage relation,
  \end{compactitem}
  \State {\bf initially} $\worklist = \set{\rootNode}$ and $N=E=U=\lhd=\emptyset$.
\end{algorithmic}
\begin{algorithmic}[1]   
  \While{$\worklist \neq \emptyset$}
  \label{ln:impact-while}

  \State dequeue $n$ from $\worklist$ 
  \label{ln:impact-dequeue}

  \State $N \leftarrow N \cup \set{n}$

  \State let $\alpha(n) = a_1, \ldots, a_k$ be the input event sequence
  on the path from $\rootNode$ to $n$

  \If{$\substform{\alpha}(\stamp{X}{1}, \ldots, \stamp{X}{k})$ is satisfiable} 
  \Comment{counterexample is feasible}

  \State get model $\nu$ of $\substform{\alpha}(\stamp{X}{1}, \ldots, \stamp{X}{k})$

  \State {\bf return} $w = (a_1,\nu(\stamp{X}{1})) \ldots (a_k,\nu(\stamp{X}{k}))$
  \label{ln:impact-nonempty}
  \Comment{$w \in L(\A)$ by construction}

  \Else \Comment{spurious counterexample}

  \State let $(I_0,\ldots,I_k)$ be a GLI for $\alpha$
  \label{ln:refine-begin}

  \State $b \leftarrow \bot$

  \For{$i=0,\ldots,k$}

  \If{$U(n_i) \not\models I_i$}

  \State $\lhd \leftarrow \lhd \setminus \set{ m \in N \mid (m,n_i) \in \lhd }$
  \Comment{uncover the nodes covered by $n_i$}

  \State $U(n_i) \leftarrow U(n_i) \wedge J_i$
  \Comment{strenghten the label of $n_i$ (Lemma \ref{lemma:refinement})}

  \If{$\neg b$}

  \State $b \leftarrow \Call{Close}{n_i}$
  \label{ln:refine-end}

  \EndIf % $\neg b$

  \EndIf % $\Lambda(n_i) \not\models I_i$

  \EndFor

  \EndIf % $\mu(n)$ is satisfiable

  \If{$n$ is not covered}

  \For{$a \in \Sigma$} 
  \Comment{expand $n$}
  \label{ln:expand-begin}

  \State let $s$ be a fresh node and $e = (n,a,s)$ be a new edge

  \State $E \leftarrow E \cup \set{e}$
  \label{ln:edge-insert}

  \State $U \leftarrow U \cup \set{(s,\top)}$

  \State enqueue $s$ into $\worklist$
  \label{ln:expand-end}

  \EndFor

  \EndIf % $n$ is not covered

  \EndWhile

  \State {\bf return} $\top$
\end{algorithmic}
\caption{\impact-based Semi-algorithm for First Order Alternating Automata}
\label{alg:impact}
\end{algorithm}

\begin{algorithm}[htb]
\begin{algorithmic}[1]
\Function{Close}{$x$} {\bf returns} $\Bool$

\For{$y \in N$ such that $\alpha(y) \prec^* \alpha(x)$}

\If{$U(x) \models U(y)$}

\State $\lhd \leftarrow \left(\lhd \setminus \set{ (p,q) \in \lhd \mid
  \text{$q$ is $x$ or a successor of $x$}}\right) \cup \set{(x,y)}$
\label{ln:close-uncover}

\State {\bf return} $\top$

\EndIf

\EndFor

\State {\bf return} $\bot$

\EndFunction
\end{algorithmic}
\caption{The \textsc{Close} function from Algorithm \ref{alg:impact}}
\end{algorithm}

Let $n$ be a node removed from the worklist at line
\ref{ln:impact-dequeue} and let $\alpha(n)$ be the input sequence
labeling the path from the root node to $n$. If
$\substform{\alpha(n)}$ is satisfiable, the sequence $\alpha(n)$ is
feasible, in which case a model of $\substform{\alpha(n)}$ is obtained
and a word $w \in L(\A)$ is returned. Otherwise, $\alpha(n)$ is an
infeasible input sequence and the procedure enters the refinement
phase (lines \ref{ln:refine-begin}-\ref{ln:refine-end}). The GLI for
$\alpha(n)$ is used to strenghten the labels of all the ancestors of
$n$, by conjoining the formulae of the interpolant, changed according
to Lemma \ref{lemma:refinement}, to the existing labels.

In this process, the nodes on the path between $\rootNode$ and $n$,
including $n$, might become eligible for coverage, therefore we
attempt to close each ancestor of $n$ that is impacted by the
refinement (line \ref{ln:refine-end}). Observe that, in this case the
call to $\Call{Close}{}$ must uncover each node which is covered by a
successor of $n$ (line \ref{ln:close-uncover} of the $\Call{Close}{}$
function). This is required because, due to the over-approximation of
the sets of reachable configurations, the covering relation is not
transitive, as explained in \cite{McMillan06}. If $\Call{Close}{}$
adds a covering edge $(n_i,m)$ to $\lhd$, it does not have to be
called for the successors of $n_i$ on this path, which is handled via
the boolean flag $b$. Finally, if $n$ is still uncovered (it has not
been previously covered during the refinement phase) we expand $n$
(lines \ref{ln:expand-begin}-\ref{ln:expand-end}) by creating a new
node for each successor $s$ via the input event $a \in \Sigma$ and
inserting it into the worklist.

\section{Proofs of the Results in the Paper}
\label{app:proofs}