### 2019-03-24 quantifiers/rebuttal.txt
ok

We first wish to thank all reviewers for their time and great
work. Below we answer the concerns raised by Reviewers 1 and 4.

Review 1:

 1.1 Relation of FOAA with s-AFA.

 Symbolic (alternating) finite-state automata are an important
 extension of finite-state automata to infinite alphabets and lacking
 those references in our paper is a minus that should be
 corrected. One explanation is that we focused on infinite alphabet
 extensions that follow an idea of the seminal papers on timed (TA)
 [1] and finite-memory (register) automata (RA) [13]. Unlike symbolic
 automata that test only the current input, both TA and RA relate
 values occurring at different positions in the input, either by using
 clocks (with reset and difference guards) or registers (with
 over-writing and equality guards). The price to be paid for this
 expressivity is that language inclusion becomes undecidable (unless
 severe restrictions on the number of clocks/registers apply).

 Our original motivation here was not as much finding decidable
 infinite alphabet models but rather having a very general model,
 (unsurprisingly) closed under complement, union and intersection, in
 which all undecidable inclusion problems can be encoded, then devise
 a practical semi-algorithm for it (that works on reasonably complex
 examples). Since, to the best of our knowledge, there are no
 published semi-algorithms for TA or RA inclusion (tools such as UPAAL
 check only for emptiness), our method could contribute to these areas
 and to program verification (we have considered several programs
 handling integer arrays in our examples).

 1.2 Restrictions of FOAA with decidable emptiness problems.

 As proved in [5], emptiness of predicate automata is undecidable even
 for quantifier-free rules with equalities and disequalities, not
 using the input variables, when predicates have arities of two or
 more (having only equalities and monadic predicates yields a
 decidable model though). In the light of this result (mentioned on
 pg. 9, last paragraph in the paper), any reasonably interesting
 extension of monadic predicate automata seems to be undecidable for
 emptiness.

 We also compare with an alternating automaton model [11], in which
 states are propositional variables and comparisons are possible only
 between adjacent input values, without hidden variables. This model
 is intermediate between s-AFA and FOAA, nevertheless emptiness
 remains undecidable, by reduction from 2-counter machine
 reachability.

Review 4:

4.1 Relation with Constrained Horn Clauses (CHC).

 The main difference between the unfoldings of CHC and those of FOAA
 is that, in CHC, the different branches of the unfolding tree are
 essentially unrelated, whereas in FOAA, all branches of the
 computation tree must synchronize on the same input word. To account
 for this synchronization, a CHC encoding of FOADA must use recursive
 algebraic data types (ADT), namely lists, to ensure that, at each
 step, all incoming branches consider exactly the same list, encoding
 the continuation of the input from that point on (CHC traverse the
 input from the end to the beginning). An encoding of FOAA in CHC
 would benefit from the advantage of using off-the-shelf CHC solvers,
 but, in our opinion, would face several problems (listed in
 increasing order of difficulty):

I. Checking that, in each clause, all incoming branches read the same
input can be done using equalities between the list arguments of the
subgoals. However, the body (left-hand side) of the clause must be
given either as a conjunction or, if quantifiers are considered, as
prenex DNF. This issue can be solved, but the reduction is a bit more
complex than suggested and, moreover, writing each rule in DNF can be
costly.

II. Introducing a list ADT in the encoding requires the use of an
 interpolation procedure for ADTs in the emptiness
 semi-algorithm. Although satisfiability of quantifier-free ADTs is
 mature and integrated with many SMT solvers, interpolation of ADT
 theories is at its beginnings. In fact, the technical report [*]
 claims to introduce the first interpolation procedure for ADTs:

* Hossein Hojjat, Philipp Ruemmer. Deciding and Interpolating
Algebraic Data Types by Reduction. ArXiv:1801.02367. Proc. of 
SYNASC'17.

Moreover, the availability of such procedures in off-the-shelf CHC
solvers is still an unresolved issue. For instance, the integration of
the above procedure in ELDARICA is underway, according to
[*]. However, when such technical problems will be solved, it would
definitely be interesting to have an experimental comparison between,
e.g. FOAA and ELDARICA.

III. Working with quantified rules in FOADA requires a quantifier
elimination procedure that produces witness terms. However, if
quantified variables occur in the ADT constructors, the theory is most
likely undecidable, thus cannot have quantifier elimination, by an
argument similar to the one of [8].

Essentially, for the above reasons, we have chosen to design our
emptiness checking semi-algorithm from scratch, trying to use as few
assumptions as possible on the underlying first-order data theory. In
fact, we only assume (a) Lyndon interpolation for quantifier-free
formulae with uninterpreted predicates and (b) witness term generation
during quantifier elimination.