### 2019-05-21 quantifiers/cav2019/body.tex

\section{Introduction}

Many results in automata theory rely on the finite alphabet
hypothesis, which guarantees, in some cases, the existence of
determinization, complementation and inclusion checking
methods. However, this hypothesis prevents the use of automata as
models of real-time systems or even simple programs, whose input and
output are data values ranging over very large domains, typically
viewed as infinite mathematical abstractions.

Traditional attempts to generalize classical Rabin-Scott automata to
infinite alphabets, such as timed automata \cite{AlurDill94} and
finite-memory automata \cite{KaminskiFrancez94} face the
\emph{complement closure} problem: there exist automata for which the
complement language cannot be recognized by an automaton in the same
class. This makes it impossible to encode a language inclusion
problem $\lang{A} \subseteq \lang{B}$ as the emptiness of an automaton
recognizing the language $\lang{A} \cap \clang{B}$, where $\clang{B}$
denotes the complement of $\lang{B}$.

Even for finite alphabets, complementation of finite-state automata
faces inherent exponential blowup, due to nondeterminism. However, if
we allow universal nondeterminism, in addition to the classical
existential nondeterminism, complementation is possible is linear
time. Having both existential and universal nondeterminism defines the
\emph{alternating automata} model \cite{ChandraKozenStockmeyer81}. A
finite-alphabet alternating automaton is described by a set of
transition rules $q \arrow{a}{} \phi$, where $q$ is a state, $a$ is an
input symbol and $\phi$ is a boolean formula, whose propositional
variables denote successor states.

\paragraph{\em Our Contribution}
We extend alternating automata to infinite data alphabets, by defining
a model of computation in which all boolean operations, including
complementation, can be done in linear time.  The control states are
given by $k$-ary predicate symbols $q(y_1,\ldots,y_k)$, the input
consists of an event $a$ from a finite alphabet and a tuple of data
variables $x_1,\ldots,x_n$, ranging over an infinite domain, and
transitions are of the form $q(y_1,\ldots,y_k)
\arrow{a(x_1,\ldots,x_n)}{} \phi(x_1, \ldots, x_n, y_1, \ldots, y_k)$,
where $\phi$ is a formula in the first-order theory of the data
domain. In this model, the arguments of a predicate atom
$q(y_1,\ldots,y_k)$ represent the values of the \emph{internal
  variables} associated with the state. Together with the input values
$x_1,\ldots,x_n$, these values define the next configurations, but
remain invisible in the input sequence.

The tight coupling of internal values and control states, by means of
uninterpreted predicate symbols, allows for linear-time
complementation just as in the case of classical propositional
alternating automata. Complementation is, moreover, possible when the
transition formulae contain first-order quantifiers, generating
infinitely-branching execution trees. The price to be paid for this
expressivity is that emptiness of first-order alternating automata is
undecidable, even for the simplest data theory of equality
\cite{Farzan15}.

The main contribution of this paper is an effective emptiness checking
semi-algorithm for first-order alternating automata, in the spirit of
the \impact~lazy annotation procedure, originally developed for
checking safety of nondeterministic integer programs
\cite{McMillan06,McMillan14}. In a nutshell, a lazy annotation
procedure unfolds an automaton $A$ trying to find an execution that
recognizes a word from $\lang{A}$. If a path that reaches a final
state does not correspond to a concrete run of the automaton, the
positions on the path are labeled with interpolants from the proof of
infeasibility, thus marking this path and all continuations as
infeasible for future searches. Termination of lazy annotation
procedures is not guaranteed, but having a suitable coverage relation
between the nodes of the search tree may ensure convergence of many
real-life examples. However, applying lazy annotation to first-order
alternating automata faces two nontrivial
problems: \begin{compactenum}
\item Quantified transition rules make it hard, if not impossible, in
  general, to decide if a path is infeasible. This is mainly because
  adding uninterpreted predicate symbols to decidable first-order
  theories, such as Presburger arithmetic, results in undecidability
  \cite{Halpern91}. To deal with this problem, we assume that the
  first-order data theory, without uninterpreted predicate symbols,
  has a quantifier elimination procedure, that instantiates quantifers
  with effectively computable \emph{witness terms}.
%
\item The interpolants that prove the infeasibility of a path are not
  \emph{local}, as they may refer to input values encountered in the
  past. However, the future executions are oblivious to \emph{when}
  these values have been seen in the past and depend only on the
  relation between the past and current values. We use this fact to
  define a labeling of nodes, visited by the lazy annotation
  procedure, with conjunctions of existentially quantified
  interpolants combining predicate atoms with data constraints.
\end{compactenum}

We use first-order alternating automata to develop practical
semi-algorithms for a number of known undecidable problems, such as:
inclusion of regular timed languages \cite{AlurDill94}, inclusion of
quasi-regular languages recognized by finite-memory automata
\cite{KaminskiFrancez94} and emptiness of predicate automata, a
subclass of first-order alternating automata used to verify
parameterized concurrent programs \cite{Farzan15,Farzan16}.

\paragraph{\em Related Work}
Recognizers for languages over infinite alphabets have found various
applications, ranging from Unicode text recognition
\cite{DAntoniKincaidWang18} to runtime program monitoring
\cite{BarringerRydeheardHavelund07}. Extending finite automata to
infinite alphabets has been considered in the context of
\emph{symbolic alternating finite automata} (s-AFA), whose transitions
are labeled with guards taken from a decidable theory of the data
domain \cite{DAntoniKincaidWang18}. As in our model, s-AFA are closed
under union, intersection and complement and emptiness is decidable,
due to the lack of registers. However, s-AFA are strictly less
expressive than our model, because comparing data at different
positions in the input word is not possible.

\emph{Constrained Horn clauses} (CHC) are a branching computation
model widespread in program verification \cite{Grebenshchikov12}. The
main difference between alternating and bottom-up branching
computations is that, in an alternating model, all branches of the
computation must synchronize on the same input word. With this in
mind, it is possible to express emptiness of first-order alternating
automata as the existence of solutions of a CHC over a higher-order
theory of data, extended with algebraic data types (lists). The
effectiveness of such an encoding depends on the effectiveness of
interpolation and witness term generation for theories of algebraic
data types \cite{HojjatRuemmer18}.

The alternating automata model presented in this paper extends the
alternating automata with variables ranging over infinite data
considered in \cite{IosifXu18}. There all variables were required to
be observable in the input. We overcome this restriction by allowing
internal (invisible) variables. Another closely related work
\cite{IosifRV16} considers an inclusion between an asynchronous
product of automata $A_1 \times \ldots \times A_n$, extended with data
variables, and a monitor automaton $B$. The semi-algorithm defined
there was based on the assumption that all variables of the observer
$B$ must be declared in the automata $A_1, \ldots, A_n$ under
check. This limitation can now be bypassed, since the inclusion
problem can be encoded as emptiness of a first-order alternating
automaton and, moreover, the emptiness checking semi-algorithm can
handle invisible variables.

The work probably closest to ours concerns the model of
\emph{predicate automata} (PA) \cite{Farzan15,Farzan16,KincaidPhD},
used in the verification of parameterized concurrent programs with
shared memory. In this model, the alphabet consists of pairs of
program statements and thread identifiers and is considered infinite
because the number of threads is unbounded. Since thread identifiers
can only be compared for equality, the data theory in PA is the theory
of equality. Even with this simplification, the emptiness problem is
undecidable when either the predicates have arity greater than one
\cite{Farzan15} or use quantified transition rules
\cite{KincaidPhD}. Checking emptiness of quantifier-free PA is
possible semi-algorithmically, by explicitly enumerating reachable
configurations and checking coverage by looking for permutations of
argument values. However, no semi-algorithm has been given for
quantified PA. Dealing with quantified transition rules is one of our
contributions.

\subsection{Preliminaries}

For two integers $0 \leq i \leq j$, we define $[i,j] \isdef
\set{i,\ldots,j}$ and $[i] \isdef [0,i]$. We consider two disjoint
sorts $\Data$ and $\Bool$, where $\Data$ is an infinite domain and
$\Bool = \set{\top,\bot}$ is the set of boolean values true ($\top$)
and false ($\bot$), respectively. The $\Data$ sort is equipped with
countably many function symbols $f : \Data^{\#(f)} \rightarrow \Data
\cup \Bool$, where $\#(f)\geq0$ denotes the number of arguments
(arity) of $f$. A \emph{predicate} is a function symbol $p :
\Data^{\#(p)} \rightarrow \Bool$ that is, a $\#(p)$-ary relation.

We consider the interpretation of all function symbols $f :
\Data^{\#(f)} \rightarrow \Data$ to be fixed by the interpretation of
the $\Data$ sort, for instance if $\Data$ is the set of integers
$\zed$, these are zero, the successor function and the arithmetic
operations of addition and multiplication. We extend this convention
to several predicates over $\Data$, such as the inequality relation
over $\zed$, and write $\preds$ for the set of remaining
\emph{uninterpreted predicates}.

Let $\vars = \set{x,y,z,\ldots}$ be a countably infinite set of
variables, ranging over $\Data$. Terms are either constants of sort
$\Data$, variables or function applications $f(t_1,\ldots,t_{\#(f)})$,
where $t_1,\ldots,t_{\#(f)}$ are terms. The set of first-order
formulae is defined by the syntax below:
\[\phi := t = s \mid p(t_1,\ldots,t_{\#(p)}) 
\mid \neg \phi_1 \mid \phi_1 \wedge \phi_2 \mid \exists x ~.~
\phi_1 \] where $t,s,t_1,\ldots,t_{\#(p)}$ denote terms and $p$ is a
predicate symbol. We write $\phi_1 \vee \phi_2$, $\phi_1 \rightarrow
\phi_2$ and $\forall x ~.~ \phi_1$ for $\neg(\neg\phi_1 \wedge
\neg\phi_2)$, $\neg\phi_1 \vee \phi_2$ and $\neg\exists x ~.~
\neg\phi_1$, respectively. $\fv{}{\phi}$ is the set of free variables
in $\phi$ and the size $\len{\phi}$ of a formula $\phi$ is the number
of symbols needed to write it down. A \emph{sentence} is a formula
$\phi$ with no free variables. A formula is \emph{positive} if each
uninterpreted predicate symbol occurs under an even number of
negations and we denote by $\posforms(Q,X)$ the set of positive
formulae with predicates from the set $Q \subseteq \preds$ and free
variables from the set $X \subseteq \vars$. A formula is in
\emph{prenex form} if it is of the form $\varphi = Q_1x_1 \ldots
Q_nx_n ~.~ \phi$, where $\phi$ has no quantifiers. In this case we
call $\phi$ the \emph{matrix} of $\varphi$. Every first-order formula
can be written in prenex form, by renaming each quantified variable to
a unique name and moving the quantifiers upfront.

%% For a variable $x \in \fv{}{\phi}$ and a term $t$, let $\phi[t/x]$ be
%% the result of replacing each occurrence of $x$ by $t$. For indexed
%% sets of terms $T=\set{t_1,\ldots,t_n}$ and variables
%% $X=\set{x_1,\ldots,x_n}$, we write $\phi[T/X]$ for the formula
%% obtained by simultaneously replacing each occurrence of $x_i$ with
%% $t_i$ in $\phi$, for all $i\in[n]$.

An \emph{interpretation} $\I$ maps each predicate symbol $p$ into a
set $p^\I \subseteq \Data^{\#(p)}$, if $\#(p)>0$, or into an element
of $\Bool$ if $\#(p)=0$. A \emph{valuation} $\nu$ maps each variable
$x$ into an element of $\Data$. Given a term $t$, we denote by $t^\nu$
the value obtained by replacing each variable $x$ by the value
$\nu(x)$ and evaluating each function application. For a formula
$\phi$, we define the forcing relation $\I,\nu \models \phi$
recursively on the structure of $\phi$, as usual. For a formula
$\phi$ and a valuation $\nu$, we define $\sem{\phi}_\nu \isdef \set{\I
  \mid \I,\nu \models \phi}$ and drop the $\nu$ subscript for
sentences. A sentence $\phi$ is \emph{satisfiable} if
$\sem{\phi}\neq\emptyset$. An element of $\sem{\phi}$ is called a
\emph{model} of $\phi$. A formula $\phi$ is \emph{valid} if $\I,\nu
\models \phi$ for every interpretation $\I$ and every valuation
$\nu$. We say that $\phi$ \emph{entails} $\psi$, written $\phi \models
\psi$ if and only if $\sem{\phi} \subseteq \sem{\psi}$.

Interpretations are partially ordered by the pointwise subset order,
defined as $\I_1 \subseteq \I_2$ if and only if $p^{\I_1} \subseteq
p^{\I_2}$ for each predicate symbol $p \in \preds$. Given a formula
$\phi$ and a valuation $\nu$, we define $\minsem{\phi}_\nu \isdef
\set{\I \mid \I,\nu \models \phi,~ \forall \I' \subseteq \I ~.~
  \I',\nu \not\models \phi}$ the set of minimal interpretations that,
together with $\nu$, form models of $\phi$.

\section{First Order Alternating Automata}

Let $\Sigma$ be a finite alphabet $\Sigma$ of \emph{input
  events}. Given a finite set of variables $X \subseteq \vars$, we
denote by $X \mapsto \Data$ the set of valuations of the variables $X$
and $\Sigma[X] = \Sigma \times (X \mapsto \Data)$ be the possibly
infinite set of \emph{data symbols} $(a,\nu)$, where $a$ is an input
symbol and $\nu$ is a valuation. A \emph{data word} (simply called
word in the following) is a finite sequence $w=(a_1,\nu_1)(a_2,\nu_2)
\ldots (a_n,\nu_n)$ of data symbols. Given a word $w$, we denote by
$\event{w} \isdef a_1\ldots a_n$ its sequence of input events and by
$\data{w}$ the valuation associating each time-stamped variable
$\stamp{x}{i}$, where $x \in \vars$, the value $\nu_i(x)$, for all
$i\in[1,n]$. We denote by $\varepsilon$ the empty sequence, by
$\Sigma^*$ the set of finite input sequences and by $\Sigma[X]^*$ the
set of finite data words over the variables $X$.

A \emph{first-order alternating automaton} is a tuple $\A =
\tuple{\Sigma,X,Q,\iota,F,\Delta}$, where $\Sigma$ is a finite set of
input events, $X$ is a finite set of input variables, $Q$ is a finite
set of predicates denoting control states, $\iota \in
\posforms(Q,\emptyset)$ is a sentence defining initial configurations,
$F \subseteq Q$ is the set of predicates denoting final states and
$\Delta$ is a set of \emph{transition rules}. A transition rule is of
the form \(q(y_1,\ldots,y_{\#(q)}) \arrow{a(X)}{} \psi\), where $q \in
Q$ is a predicate, $a \in \Sigma$ is an input event and $\psi \in
\posforms(Q,X \cup \set{y_1,\ldots,y_{\#(q)}})$ is a positive formula,
where $X \cap \set{y_1,\ldots,y_{\#(q)}} = \emptyset$.  Without loss
of generality, we consider, for each predicate $q \in Q$ and each
input event $a \in \Sigma$, at most one such rule, as two or more
rules can be joined using disjunction. The quantifiers occurring in
the right-hand side formula of a transition rule are called
\emph{transition quantifiers}. The \emph{size} of $\A$ is $\len{\A}
\isdef \len{\iota} + \sum \set{\len{\psi} \mid q(\vec{y})
  \arrow{a(X)}{} \psi \in \Delta}$.

The semantics of first-order alternating automata is analogous to the
semantics of propositional alternating automata, with rules of the
form $q \arrow{a}{} \phi$, where $q$ is a propositional variable and
$\phi$ a positive boolean combination of propositional variables. For
instance, $q_0 \arrow{a}{} (q_1 \wedge q_2) \vee q_3$ means that the
automaton can choose to transition in either both $q_1$ and $q_2$ or
in $q_3$ alone. This leads to defining transitions as the
\emph{minimal models} of the right hand side of a rule\footnote{Both
  $\set{q_1 \leftarrow \top, q_2 \leftarrow \top, q_3 \leftarrow
    \bot}$ and $\set{q_1 \leftarrow \bot, q_2 \leftarrow \bot, q_3
    \leftarrow \top}$ are minimal models, however $\set{q_1 \leftarrow
    \top, q_2 \leftarrow \top, q_3 \leftarrow \top}$ is a model but is
  not minimal.}. The original definition of alternating automata
\cite{ChandraKozenStockmeyer81} works around this problem and
considers boolean valuations instead of formulae. In contrast, a
finite description of a first-order alternating automaton cannot be
given in terms of interpretations, as a first-order formula may have
infinitely many models, corresponding to infinitely many initial or
successor states occurring within an execution step.

Given an uninterpreted predicate symbol $q \in Q$ and data
values $d_1,\ldots,d_{\#(q)} \in \Data$, the tuple $(q, d_1,\ldots,d_{\#(q)})$
is called a \emph{configuration}, sometimes written
$q(d_1,\ldots,d_{\#(q)})$, when no confusion arises. A configuration
is \emph{final} if $q \in F$. An interpretation $\I$ corresponds to a
set of configurations $\cube{\I} \isdef \set{(q, d_1,\ldots,d_{\#(q)})
  \mid q \in Q,~ (d_1,\ldots,d_{\#(q)}) \in q^\I}$, called a
\emph{cube}. This notation is lifted to sets of configurations in the
usual way.

\begin{definition}\label{def:execution}
Given a word $w=(a_1,\nu_1) \ldots (a_n,\nu_n) \in \Sigma[X]^*$ and a
cube $c$, an \emph{execution} of
$\A=\tuple{\Sigma,X,Q,\iota,F,\Delta}$ over $w$, starting with $c$, is
a forest $\T = \set{T_1,T_2,\ldots}$, where each $T_i$ is a tree
labeled with configurations, such that:
\begin{compactenum}
\item\label{it1:execution} $c = \set{T(\epsilon) \mid T \in \T}$ is the set of
  configurations labeling the roots of $T_1,T_2,\ldots$ and
%
\item\label{it2:execution} if $(q,d_1,\ldots,d_{\#(q)})$ labels a node
  on the level $j \in [n-1]$ in $T_i$, then the labels of its children
  form a cube from $\cube{\minsem{\psi}_{\eta}}$, where $\eta =
  \nu_{j+1}[y_1 \leftarrow d_1,\ldots,y_{\#(q)} \leftarrow d_{\#(q)}]$
  and \(q(y_1,\ldots,y_{\#(q)}) \arrow{a_{j+1}(X)}{} \psi \in \Delta\)
  is a transition rule of $\A$.
\end{compactenum}
\end{definition}

An execution $\T$ over $w$, starting with $c$, is \emph{accepting} if
and only if all paths in $\T$ have the same length and the frontier of
each tree $T \in \T$ is labeled with final configurations.  If $\A$
has an accepting execution over $w$ starting with a cube $c \in
\cube{\minsem{\iota}}$, then $\A$ \emph{accepts} $w$ and let
$\lang{\A}$ be the set of words accepted by $\A$. For example,
consider the automaton $\A = \tuple{\set{a}, \set{x},
  \set{q_0,q_1,q_2,q_f}, q_0(0), \set{q_f}, \Delta}$, where $\Delta$
is the set: \(q_0(y) \arrow{a(x)}{} q_1(y+x) \wedge q_2(y-x)\),
\(q_1(y) \arrow{a(x)}{} q_1(y+x) \vee (y > 0 \wedge q_f)\) and
\(q_2(y) \arrow{a(x)}{} q_2(y-x) \vee (y > 0 \wedge q_f)\). A possible
execution tree of this automaton is the following:
\begin{center}
  \input exec.pdf_t
\end{center}
The execution tree is not accepting, since its frontier is not
labeled with final configurations everywhere. Incidentally, here we
have $\lang{\A} = \emptyset$, which is proved by our tool in
$\sim\!0.5$ seconds on an average machine.

\noindent
In the rest of this paper, we are concerned with the following
problems: \begin{compactenum}
\item \emph{boolean closure}: given automata $\A_i =
  \tuple{\Sigma,X,Q_i,\iota_i,F_i,\Delta_i}$, for $i=1,2$, do there
  exist automata $\A_\cap$, $\A_\cup$ and $\overline{\A}_1$ such that
  $L(\A_\cap) = L(\A_1) \cap L(\A_2)$, $L(\A_\cup) = L(\A_1) \cup
  L(\A_2)$ and $L(\overline{\A}_1) = \Sigma[X]^* \setminus L(\A_1)$ ?
%
\item \emph{emptiness}: given an automaton $\A$, is $L(\A) =
  \emptyset$ ?
\end{compactenum}
For technical reasons, we address the following problem next: given an
automaton $\A$ and an input sequence $\alpha \in \Sigma^*$, does there
exists a word $w \in \lang{\A}$ such that $\event{w} = \alpha$ ?  By
solving this problem first, we develop the machinery required to prove
that first-order alternating automata are closed under complement and,
further, set up the ground for developping a practical semi-algorithm
for the emptiness problem.

\subsection{Path Formulae}
\label{sec:path-formulae}

In the upcoming developments it is sometimes more convenient to work
with logical formulae defining executions of automata, than with
low-level execution forests. For this reason, we first introduce
\emph{path formulae} $\pathform{\alpha}$, which are formulae defining
the executions of an automaton, over words that share a given sequence
$\alpha$ of input events. Second, we restrict a path formula
$\pathform{\alpha}$ to an \emph{acceptance formula}
$\accform{\alpha}$, which defines only those executions that are
accepting among $\pathform{\alpha}$. Consequently, the automaton
accepts a word $w$ such that $\event{w} = \alpha$ if and only if
$\accform{\alpha}$ is satisfiable.

Let $\A = \tuple{\Sigma,X,Q,\iota,F,\Delta}$ be an automaton for the
rest of this section. For any $i \in \nat$, we denote by $\stamp{Q}{i}
= \set{\stamp{q}{i} \mid q \in Q}$ and $\stamp{X}{i} =
\set{\stamp{x}{i} \mid x \in X}$ the sets of time-stamped predicate
symbols and variables, respectively. We also define $\stamp{Q}{\leq n}
\isdef \set{\stamp{q}{i} \mid q \in Q, i \in [n]}$ and $\stamp{X}{\leq
  n} \isdef \set{\stamp{x}{i} \mid x \in X, i \in [n]}$. For a formula
$\psi$ and $i \in \nat$, we define $\stamp{\psi}{i} \isdef
\psi[\stamp{X}{i}/X,\stamp{Q}{i}/Q]$ the formula in which all input
variables and state predicates (and only those symbols) are replaced
by their time-stamped counterparts. Moreover, we write $q(\vec{y})$
for $q(y_1,\ldots,y_{\#(q)})$, when no confusion arises.

Given a sequence of input events $\alpha = a_1 \ldots a_n \in
\Sigma^*$, the \emph{path formula} of $\alpha$ is:
\begin{equation}\label{eq:pathform}
\begin{array}{c}
\pathform{\alpha} \isdef \stamp{\iota}{0} \wedge 
\bigwedge_{i=1}^n \bigwedge_{q(\vec{y}) \arrow{a_i(X)}{}
  \psi \in \Delta} \forall y_1 \ldots \forall y_{\#(q)} ~.~
\stamp{q}{i-1}(\vec{y}) \rightarrow \stamp{\psi}{i}
\end{array}
\end{equation}
The automaton $\A$, to which $\pathform{\alpha}$ refers, will always
be clear from the context. To formalize the relation between the
low-level configuration-based execution semantics and path formulae,
consider a word $w=(a_1,\nu_1) \ldots (a_n,\nu_n) \in
\Sigma[X]^*$. Any execution $\T$ of $\A$ over $w$ has an associated
interpretation $\I_{\T}$ of time-stamped predicates $\stamp{Q}{\leq
  n}$:
\[\I_{\T}(\stamp{q}{i}) \isdef \set{(d_1,\ldots,d_{\#(q)}) \mid
  (q,d_1,\ldots,d_{\#(q)}) \text{ labels a node on level $i$ in $\T$}}
,~\forall q\in Q ~\forall i \in [n]\]

\begin{lemma}\label{lemma:path-formula}
  Given an automaton $\A = \tuple{\Sigma,X,Q,\iota,F,\Delta}$, for any
  word $w=(a_1,\nu_1) \ldots (a_n,\nu_n)$, we have 
  \(\minsem{\pathform{\event{w}}}_{\data{w}} = \set{\I_{\T} \mid \T
    \text{ is an execution of $\A$ over $w$}}\). 
\end{lemma}

Next, we give a logical characterization of acceptance, relative to a
given sequence of input events $\alpha \in \Sigma^*$. To this end, we
constrain the path formula $\pathform{\alpha}$ by requiring that only
final states of $\A$ occur on the last level of the execution.  The
result is the \emph{acceptance formula} for $\alpha$:
\begin{equation}\label{eq:accform}
\begin{array}{c}
\accform{\alpha} \isdef \pathform{\alpha} \wedge \bigwedge_{q
  \in Q \setminus F} \forall y_1 \ldots \forall y_{\#(q)} ~.~
\stamp{q}{n}(\vec{y}) \rightarrow \bot
\end{array}
\end{equation}
The top-level universal quantifiers from a subformula $\forall y_1
\ldots \forall y_{\#(q)} ~.~ \stamp{q}{i}(\vec{y}) \rightarrow \psi$
of $\accform{\alpha}$ will be referred to as \emph{path quantifiers},
in the following. Notice that path quantifiers are distinct from the
transition quantifiers that occur within a formula $\psi$ of a
transition rule $q(y_1,\ldots,y_{\#(q)}) \arrow{a(X)}{} \psi$ of $\A$.
The relation between the words accepted by $\A$ and the acceptance
formula above, is formally captured by the following lemma:

\begin{lemma}\label{lemma:acceptance}
  Given an automaton $\A = \tuple{\Sigma,X,Q,\iota,F,\Delta}$, for
  every word $w \in \Sigma[X]^*$, the following are
  equivalent: \begin{inparaenum}[(1)]
  \item\label{it1:lemma:acceptance} there exists an interpretation
    $\I$ such that $\I,\data{w} \models \accform{\event{w}}$ and
    %
  \item\label{it2:lemma:acceptance} $w \in \lang{\A}$. 
  \end{inparaenum}
\end{lemma}

As an immediate consequence, one can decide whether $\A$ accepts some
word $w$ with a given input sequence $\event{w}=\alpha$, by checking
whether $\accform{\alpha}$ is satisfiable. However, unlike
non-alternating infinite-state models of computation, such as counter
automata (nondeterministic programs with integer variables), the
satisfiability query for an acceptance (path) formula falls outside of
known decidable theories, supported by standard SMT solvers. There are
basically two reasons for this, namely\ \begin{inparaenum}[(i)]
\item the presence of predicate symbols, and
\item the non-trivial alternation of quantifiers.
\end{inparaenum}
To understand this point, consider for example, the decidable theory
of Presburger arithmetic \cite{Presburger29}. Adding even only one
monadic predicate symbol to it yields undecidability in the presence
of non-trivial quantifier alternation \cite{Halpern91}. On the other
hand, the quantifier-free fragment of Presburger arithmetic extended
with uninterpreted function symbols is decidable, by a Nelson-Oppen
style congruence closure argument \cite{NelsonOppen80}.

To tackle the problem of deciding satisfiability of $\accform{\alpha}$
formulae, we start from the observation that their form is rather
particular, which allows the elimination of path quantifiers and
uninterpreted predicate symbols, by a couple of
satisfiability-preserving transformations. The result of applying
these transformations is a formula with no predicate symbols, whose
only quantifiers are those introduced by the transition rules of the
automaton. Next, in \S\ref{sec:emptiness} we shall assume moreover
that the first-order theory of the data sort $\Data$ (without
uninterpreted predicate symbols) has quantifier elimination, providing
thus an effective decision procedure.

For the time being, let us formally define the elimination of
transition quantifiers and predicate symbols. Let $\alpha = a_1 \ldots
a_n$ be a given sequence of input events and let $\alpha_i$ be the
prefix $a_1 \ldots a_i$ of $\alpha$, for $i \in [n]$, where
$\alpha_0=\epsilon$. We consider the sequence of formulae
$\quantformn{\alpha_0}, \ldots, \quantformn{\alpha_n}$ defined as
$\quantformn{\alpha_0} \isdef \stamp{\iota}{0}$ and, for all $i \in
[1,n]$, let $\quantformn{\alpha_i}$ be the conjunction of
$\quantformn{\alpha_{i-1}}$ with all formulae
$\stamp{q}{i-1}(t_1,\ldots,t_{\#(q)}) \rightarrow
\stamp{\psi}{i}[t_1/y_1, \ldots, t_{\#(q)}/y_{\#(q)}]$, such that
$\stamp{q}{i-1}(t_1,\ldots,t_{\#(q)})$ occurs in
$\quantformn{\alpha_{i-1}}$, for some terms
$t_1,\ldots,t_{\#(q)}$. Next, we write $\quantform{\alpha}$ for the
conjunction of $\quantformn{\alpha_n}$ with all
$\stamp{q}{n}(t_1,\ldots,t_{\#(q)}) \rightarrow \bot$, such that
$\stamp{q}{n}(t_1,\ldots,t_{\#(q)})$ occurs in
$\quantformn{\alpha_n}$, for some $q \in Q \setminus F$. Note that
$\quantform{\alpha}$ contains no path quantifiers, as required. On the
other hand, the scope of the transition quantifiers in
$\quantform{\alpha}$ exceeds the right-hand side formulae from the
transition rules, as shown by the following example.

\begin{example}\label{ex:quant-elim}
  Consider the automaton $\A = \tuple{\set{a_1,a_2}, \set{x},
    \set{q,q_f}, \iota, \set{q_f}, \Delta}$, where:
  \[\begin{array}{rcl}
  \iota & = & \exists z ~.~ z \geq 0 \wedge q(z) \\
  \Delta & = & \set{
    q(y) \arrow{a_1(x)}{} x\geq 0 \wedge \forall z ~.~ z \leq y \rightarrow q(x+z),~
    q(y) \arrow{a_2(x)}{} y<0 \wedge q_f(x+y)
  }
  \end{array}\]
  For the input event sequence $\alpha = a_1a_2$, the acceptance
  formula is:
  \[\begin{array}{rcl}
  \accform{\alpha} & = & \exists z_1 ~.~ z_1 \geq 0 \wedge \stamp{q}{0}(z_1) ~\wedge \\
  && \forall y ~.~ \stamp{q}{0}(y) \rightarrow [\stamp{x}{1}\geq0 \wedge \forall z_2 ~.~ z_2\geq y 
    \rightarrow \stamp{q}{1}(\stamp{x}{1}+z_2)] ~\wedge \\
  && \forall y ~.~ \stamp{q}{1}(y) \rightarrow [y<0 \wedge \stamp{q_f}{2}(\stamp{x}{2}+y)]
  \end{array}\]
  The result of eliminating the path quantifiers, in prenex normal form, is shown below:
  \[\begin{array}{rcl}
  \quantform{\alpha} & = & \exists z_1\forall z_2 ~.~ z_1 \geq 0 \wedge \stamp{q}{0}(z_1) ~\wedge \\
  && [\stamp{q}{0}(z_1) \rightarrow \stamp{x}{1} \geq 0 \wedge (z_2 \geq z_1 \rightarrow \stamp{q}{1}(\stamp{x}{1}+z_2))] ~\wedge \\
  && [\stamp{q}{1}(\stamp{x}{1}+z_2) \rightarrow \stamp{x}{1}+z_2 < 0 \wedge \stamp{q_f}{2}(\stamp{x}{2}+\stamp{x}{1}+z_2)]   
  \end{array}\]
  Notice that the transition quantifiers $\exists z_1$ and $\forall
  z_2$ from $\accform{\alpha}$ range now over
  $\quantform{\alpha}$. \hfill$\blacksquare$
\end{example}

\begin{lemma}\label{lemma:quant}
  For any input event sequence $\alpha=a_1\ldots a_n$ and each
  valuation $\nu : \stamp{X}{\leq n} \rightarrow \Data$, the following
  hold, for every interpretation $\I$: \begin{inparaenum}[(1)]
  \item\label{it1:quant} if $\I,\nu \models \accform{\alpha}$ then
    $\I,\nu \models \quantform{\alpha}$, and
    %
  \item\label{it2:quant} if $\I,\nu \models \quantform{\alpha}$ there
    exists an interpretation $\J \subseteq \I$ such that $\J,\nu
    \models \accform{\alpha}$.
  \end{inparaenum}
\end{lemma}

Further, we eliminate the predicate atoms from $\quantform{\alpha}$,
by considering the sequence of formulae $\substformn{\alpha_0} \isdef
\stamp{\iota}{0}$ and $\substformn{\alpha_i}$ is obtained by
substituting each predicate atom
$\stamp{q}{i-1}(t_1,\ldots,t_{\#(q)})$ in $\substformn{\alpha_{i-1}}$
by $\stamp{\psi}{i}[t_1/y_1,\ldots,t_{\#(q)}/y_{\#(q)}]$, where
$q(\vec{y}) \arrow{a_{i}(X)}{} \psi \in \Delta$, for all $i \in
[1,n]$. We write $\substform{\alpha}$ for the formula obtained by
replacing, in $\substformn{\alpha}$, each occurrence of a predicate
$\stamp{q}{n}$, such that $q \in Q \setminus F$ (resp. $q \in F$), by
$\bot$ (resp. $\top$).
\begin{example}[Contd. from Example \ref{ex:quant-elim}]\label{ex:pred-elim}
  The result of the elimination of predicate atoms from the acceptance
  formula in Example \ref{ex:quant-elim} is shown below:
  \[\substform{\alpha} = 
  \exists z_1 \forall z_2 ~.~ z_1 \geq 0 \wedge 
          [\stamp{x}{1} \geq 0 \wedge (z_2 \geq z_1 \rightarrow
            \stamp{x}{1}+z_2 < 0)]\] Since this formula is
          unsatisfiable, by Lemma \ref{lemma:quant-pred-acceptance}
          below, no word $w$ with input event sequence $\event{w} =
          a_1a_2$ is accepted by the automaton $\A$ from Example
          \ref{ex:quant-elim}. \hfill$\blacksquare$
\end{example}

At this point, we prove the formal relation between the satisfiability
of the formulae $\quantform{\alpha}$ and $\substform{\alpha}$. Since
there are no occurrences of predicates in $\substform{\alpha}$, for
each valuation $\nu : \stamp{X}{\leq n} \rightarrow \Data$, there
exists an interpretation $\I$ such that $\I,\nu \models
\substform{\alpha}$ if and only if $\J,\nu \models
\substform{\alpha}$, for every interpretation $\J$. In this case we
omit $\I$ and simply write $\nu \models \substform{\alpha}$.

\begin{lemma}\label{lemma:subst}
  For any input event sequence $\alpha=a_1\ldots a_n$ and each
  valuation $\nu : \stamp{X}{\leq n} \rightarrow \Data$, there exists
  a valuation $\I$ such that $\I,\nu \models \quantform{\alpha}$ if
  and only if $\nu \models \substform{\alpha}$. 
\end{lemma}

Finally, we define the acceptance of a word with a given input event
sequence by means of a quantifier-free formula in which no predicate
atom occurs.

\begin{lemma}\label{lemma:quant-pred-acceptance}
  Given an automaton $\A = \tuple{\Sigma,X,Q,\iota,F,\Delta}$, for
  every word $w \in \Sigma[X]^*$, we have $\data{w} \models
  \substform{\event{w}}$ if and only if $w \in \lang{\A}$.
\end{lemma}

\subsection{Boolean Closure of First Order Alternating Automata}
\label{sec:closure}

Given a positive formula $\phi$, we define the \emph{dual} formula
$\dual{\phi}$ recursively as follows:
\[\begin{array}{rclcrclcrcl}
\dual{(\phi_1 \vee \phi_2)} & \isdef & \dual{\phi_1} \wedge \dual{\phi_2} && 
\dual{(\phi_1 \wedge \phi_2)} & \isdef & \dual{\phi_1} \vee \dual{\phi_2} &&
\dual{(t = s)} & \isdef & t \neq s \\
\dual{(\exists x ~.~ \phi_1)} & \isdef & \forall x ~.~ \dual{\phi_1} && 
\dual{(\forall x ~.~ \phi_1)} & \isdef & \exists x ~.~ \dual{\phi_1} && 
\dual{(t \neq s)} & \isdef & t = s \\
&&&& \dual{q(x_1,\ldots,x_{\#(q)})} & \isdef & q(x_1,\ldots,x_{\#(q)})
\end{array}\]
The following theorem shows closure of automata under all boolean
operations. Note that it is sufficient to show closure under
intersection and negation because $\lang{\A_1} \cup \lang{\A_2}$ is
the complement of the language $\clang{\A_1} \cap \clang{\A_2}$, for
any two automata $\A_1$ and $\A_2$ with the same input event alphabet
and set of input variables.

\begin{theorem}\label{thm:closure}
  Given automata $\A_i = \tuple{\Sigma,X,Q_i,\iota_i,F_i,\Delta_i}$,
  for $i=1,2$, such that $Q_1 \cap Q_2 = \emptyset$, the following hold: 
  \begin{compactenum}
    \item\label{it1:thm:closure} $\lang{\A_\cap} = \lang{\A_1} \cap \lang{\A_2}$, where
      $\A_\cap = \tuple{\Sigma,X,Q_1 \cup Q_2, \iota_1 \wedge \iota_2,
      F_1 \cup F_2, \Delta_1 \cup \Delta_2}$, 
      %
    \item\label{it2:thm:closure} $\lang{\overline{\A_i}} = \Sigma[X]^*
      \setminus \lang{\A_i}$, where $\overline{\A_i} =
      \tuple{\Sigma,X,Q_i,\dual{\iota},Q_i\setminus
        F_i,\dual{\Delta}_i}$ and \(\dual{\Delta}_i =
      \set{q(\vec{y}) \arrow{a(X)}{} \dual{\psi} \mid
        q(\vec{y}) \arrow{a(X)}{} \psi \in
        \Delta_i}\), for $i=1,2$. 
  \end{compactenum}
  Moreover, $\len{\A_\cap} = \bigO{\len{\A_1}+\len{\A_2}}$ and
  $\len{\overline{\A_i}} = \bigO{\len{\A_i}}$, for $i=1,2$.
\end{theorem}

\section{The Emptiness Problem}
\label{sec:emptiness}

The emptiness problem is undecidable even for automata with predicates
of arity two, whose transition rules use only equalities and
disequalities, having no transition quantifiers \cite{Farzan15}. Since
even such simple classes of alternating automata have no general
decision procedure for emptiness, we use an abstraction-refinement
semi-algorithm based on \emph{lazy annotation}
\cite{McMillan06,McMillan14}. In a nutshell, a lazy annotation
procedure systematically explores the set of finite input event
sequences searching for an accepting execution. For an input sequence,
if the path formula is satisfiable, we compute a word in the language
of the automaton, from the model of the path formula. Otherwise,
i.e.\ the sequence is \emph{spurious}, the search backtracks and each
position in the sequence is annotated with an interpolant, thus
marking the sequence as infeasible. The semi-algorithm uses moreover a
coverage relation between sequences, ensuring that the continuations
of already covered sequences are never explored. Sometimes this
coverage relation provides a sound termination argument, in case when
the automaton is empty.

For two input event sequences $\alpha, \beta \in \Sigma^*$, we say
that $\alpha$ is a prefix of $\beta$, written $\alpha \prefix \beta$,
if $\alpha=\beta\gamma$ for some sequence $\gamma\in\Sigma^*$. A set
$S$ of sequences is \emph{prefix-closed} if for each $\alpha \in S$,
if $\beta \prefix \alpha$ then $\beta \in S$, and \emph{complete} if
for each $\alpha \in S$, there exists $a \in \Sigma$ such that $\alpha
a \in S$ if and only if $\alpha b \in S$ for all $b \in \Sigma$. A
prefix-closed set is the backbone of a tree whose edges are labeled
with input events. If the set is, moreover, complete, then every node
of the tree has either zero successors, in which case it is called a
\emph{leaf}, or it has a successor edge labeled with $a$ for each
input event $a \in \Sigma$.

\begin{definition}\label{def:unfolding}
  An \emph{unfolding} of an automaton $\A =
  \tuple{\Sigma,X,Q,\iota,F,\Delta}$ is a finite partial mapping $U :
  \Sigma^* \rightharpoonup_{\mathit{fin}} \posforms(Q,\emptyset)$,
  whose domain $\dom(U)$ is a finite prefix-closed complete set, such
  that $U(\epsilon) = \iota$, and for each sequence $\alpha a \in
  \dom(U)$, such that $\alpha \in \Sigma^*$ and $a \in \Sigma$:
  \[\begin{array}{c}
  \stamp{U(\alpha)}{0} \wedge
  \bigwedge_{q(\vec{y}) \arrow{a(X)}{} \psi} \forall y_1 \ldots
  \forall y_{\#q} ~.~ \stamp{q}{0}(\vec{y}) \rightarrow
  \stamp{\psi}{1} \models \stamp{U(\alpha a)}{1}
  \end{array}\]
  A path $\alpha$ is \emph{safe} in $U$ if and only if $U(\alpha)
  \wedge \bigwedge_{q \in Q \setminus F} \forall y_1 \ldots
  \forall_{y_{\#(q)}} ~.~ q(\vec{y}) \rightarrow \bot$ is
  unsatisfiable. The unfolding $U$ is safe if and only if every path
  in $\dom(U)$ is safe in $U$.
\end{definition}

Lazy annotation semi-algorithms \cite{McMillan06,McMillan14} build
unfoldings of automata trying to discover counterexamples for
emptiness. If the automaton $\A$ in question is non-empty, a
systematic enumeration of the input event sequences\footnote{For
  instance, using breadth-first search.} from $\Sigma^*$ will suffice
to discover a word $w \in \lang{\A}$, provided that the first-order
theory of the data domain $\Data$ is decidable (Lemma
\ref{lemma:acceptance}). However, if $\lang{\A} = \emptyset$, the
enumeration of input event sequences may, in principle, run
forever. The typical way of fighting this divergence problem is to
define a \emph{coverage} relation between the nodes of the unfolding
tree.

\begin{definition}\label{def:coverage}
  Given an unfolding $U$ of an automaton $\A =
  \tuple{\Sigma,X,Q,\iota,F,\Delta}$ a node $\alpha \in \dom(U)$ is
  \emph{covered} by another node $\beta \in \dom(U)$, denoted $\alpha
  \cover \beta$, if and only if there exists a node $\alpha' \prefix
  \alpha$ such that $U(\alpha') \models U(\beta)$. Moreover, $U$ is
  \emph{closed} if and only if every leaf from $\dom(U)$ is covered by
  an uncovered node.
\end{definition}

\input algorithm

A lazy annotation semi-algorithm will stop and report emptiness
provided that it succeeds in building a closed and safe unfolding of
the automaton. Notice that, by Definition \ref{def:coverage}, for any
three nodes of an unfolding $U$, say $\alpha,\beta,\gamma \in
\dom(U)$, if $\alpha \prec \beta$ and $\alpha \cover \gamma$, then
$\beta \cover \gamma$ as well. As we show next (Theorem
\ref{thm:soundness}), there is no need to expand covered nodes,
because, intuitively, there exists a word $w \in \lang{\A}$ such that
$\alpha \preceq \event{w}$ and $\alpha \cover \gamma$ only if there
exists another word $u \in \lang{\A}$ such that $\gamma \preceq
\event{u}$. Hence, exploring only those input event sequences that are
continuations of $\gamma$ (and ignoring those of $\alpha$) suffices in
order to find a counterexample for emptiness, if one exists.

An unfolding node $\alpha \in \dom(U)$ is said to be \emph{spurious}
if and only if $\accform{\alpha}$ is unsatisfiable. In this case, we
change (refine) the labels of (some of the) prefixes of $\alpha$ (and
that of $\alpha$), such that $U(\alpha)$ becomes $\bot$, thus
indicating that there is no real execution of the automaton along that
input event sequence. As a result of the change of labels, if a node
$\gamma \prefix \alpha$ used to cover another node from $\dom(U)$, it
might not cover it with the new label. Therefore, the coverage
relation has to be recomputed after each refinement of the
labeling. The semi-algorithm stops when (and if) a safe complete
unfolding has been found.

\begin{theorem}\label{thm:soundness}
  If an automaton $\A$ has a nonempty safe closed unfolding then
  $\lang{A} = \emptyset$.
\end{theorem}

We describe the semi-algorithm used to check emptiness of first-order
alternating automata. The execution of Algorithm \ref{alg:impact}
consists of three phases, corresponding to the \textsc{Close},
\textsc{Refine} and \textsc{Expand} of the original \impact\ procedure
\cite{McMillan06}.  Let $n$ be a node removed from the worklist at
line \ref{ln:impact-dequeue} and let $\alpha(n)$ be the input sequence
labeling the path from the root node to $n$. If
$\substform{\alpha(n)}$ is satisfiable, the sequence $\alpha(n)$ is
feasible, in which case a model of $\substform{\alpha(n)}$ is obtained
and a word $w \in L(\A)$ is returned. Otherwise, $\alpha(n)$ is an
infeasible input sequence and the procedure enters the refinement
phase (lines \ref{ln:refine-begin}-\ref{ln:refine-end}). The GLI for
$\alpha(n)$ is used to strenghten the labels of all the ancestors of
$n$, by conjoining the formulae of the interpolant, changed according
to Lemma \ref{lemma:refinement}, to the existing labels.

In this process, the nodes on the path between $\rootNode$ and $n$,
including $n$, might become eligible for coverage, therefore we
attempt to close each ancestor of $n$ that is impacted by the
refinement (line \ref{ln:refine-end}). Observe that, in this case the
call to $\Call{Close}{}$ must uncover each node which is covered by a
successor of $n$ (line \ref{ln:close-uncover} of the $\Call{Close}{}$
function). This is required because, due to the over-approximation of
the sets of reachable configurations, the covering relation is not
transitive, as explained in \cite{McMillan06}. If $\Call{Close}{}$
adds a covering edge $(n_i,m)$ to $\lhd$, it does not have to be
called for the successors of $n_i$ on this path, which is handled via
the boolean flag $b$. Finally, if $n$ is still uncovered (it has not
been previously covered during the refinement phase) we expand $n$
(lines \ref{ln:expand-begin}-\ref{ln:expand-end}) by creating a new
node for each successor $s$ via the input event $a \in \Sigma$ and
inserting it into the worklist.

\section{Interpolant Generation}\label{sec:interpolants}

Typically, when checking the unreachability of a set of program
configurations, the interpolants used to annotate the unfolded control
structure are assertions about the values of the program variables in
a given control state, at a certain step of an execution
\cite{McMillan06}. Because we consider alternating computation trees
(forests), we must distinguish between
\begin{inparaenum}[(i)]
\item locality of interpolants w.r.t. a given control state (control
  locality) and
%
\item locality w.r.t. a given time stamp (time locality).
\end{inparaenum} 
In logical terms, \emph{control-local} interpolants are formulae
involving a single predicate symbol, whereas \emph{time-local}
interpolants involve only predicates $\stamp{q}{i}$ and variables
$\stamp{x}{i}$, for a single $i \geq 0$. When considering alternating
executions, control-local interpolants are not always enough to prove
emptiness, because of the synchronization of several branches of the
computation on the same input word. For this reason, the interpolants
considered in this paper will never be control-local and we shall use
the term \emph{local} to denote time-local interpolants, with no free
variables.

First, let us give the formal definition of the class of interpolants
we shall work with. Given a formula $\phi$, the \emph{vocabulary} of
$\phi$, denoted $\voc{\phi}$ is the set of predicate symbols $q \in
\stamp{Q}{i}$ and variables $x \in \stamp{X}{i}$, occurring in $\phi$,
for some $i\geq0$. For a term $t$, its vocabulary $\voc{t}$ is the set
of variables that occur in $t$. Observe that quantified variables and
the interpreted function symbols of the data
theory\footnote{E.g.,\ the arithmetic operators of addition and
  multiplication, when $\Data$ is the set of integers.}  do not belong
to the vocabulary of a formula. By $\pset{+}{\phi}$ [$\pset{-}{\phi}$]
we denote the set of predicate symbols that occur in $\phi$ under an
even [odd] number of negations.

\begin{definition}[\cite{Lyndon59}]\label{def:lyndon-interpolant}
Given formulae $\phi$ and $\psi$ such that $\phi \wedge \psi$ is
unsatisfiable, a \emph{Lyndon interpolant} is a formula $I$ such that
$\phi \models I$, the formula $I \wedge \psi$ is unsatisfiable,
$\voc{I} \subseteq \voc{\phi} \cap \voc{\psi}$, $\pset{+}{I} \subseteq
\pset{+}{\phi} \cap \pset{+}{\psi}$ and $\pset{-}{I} \subseteq
\pset{-}{\phi} \cap \pset{-}{\psi}$.
\end{definition}

In the rest of this section, fix an automaton $\A =
\tuple{\Sigma,X,Q,\iota,F,\Delta}$. The following definition
generalizes interpolants from unsatisfiable conjunctions to input
sequences:

\begin{definition}\label{def:generalized-lyndon-interpolant}
  Given a sequence of input events $\alpha = a_1 \ldots a_n \in
  \Sigma^*$, a \emph{generalized Lyndon interpolant (GLI)} is a
  sequence $(I_0,\ldots,I_n)$ of formulae such that, for all $k \in
  [n-1]$, the following hold: \begin{inparaenum}[(1)]
    \item\label{it1:generalized-lyndon-interpolant} $\pset{-}{I_k} =
      \emptyset$,
      %
    \item\label{it2:generalized-lyndon-interpolant} $\stamp{\iota}{0}
      \models I_0$, \(I_k \wedge \Big(\bigwedge_{q(\vec{y})
        \arrow{a_i(X)}{} \psi \in \Delta} \forall y_1 \ldots \forall
      y_{\#(q)} ~.~ \stamp{q}{k}(\vec{y}) \rightarrow
      \stamp{\psi}{k+1}\Big) \models I_{k+1}\) and
      %
    \item\label{it3:generalized-lyndon-interpolant} $I_n \wedge
      \bigwedge_{q \in Q \setminus F} \forall y_1 \ldots \forall
      y_{\#(q)} ~.~ q(\vec{y}) \rightarrow \bot$ is unsatisfiable.
  \end{inparaenum}
  Moreover, the GLI is \emph{local} if and only if $\voc{I_k}
  \subseteq \stamp{Q}{k}$, for all $k \in [n]$.
\end{definition}
The following proposition states the existence of local GLI for the
theories in which Lyndon's Interpolation Theorem holds.

\begin{proposition}\label{prop:local-gli}
  If there exists a Lyndon interpolant for any two formulae $\phi$ and
  $\psi$, in the first-order theory of data with uninterpreted
  predicate symbols, such that $\phi \wedge \psi$ is unsatisfiable,
  then any sequence of input events $\alpha = a_1 \ldots a_n \in
  \Sigma^*$, such that $\accform{\alpha}$ is unsatisfiable, has a
  local GLI $(I_0,\ldots,I_n)$.
\end{proposition}
A problematic point of the above proposition is that the existence of
Lyndon interpolants (Definition \ref{def:lyndon-interpolant}) is
proved in principle, but the proof is non-constructive. In other
words, the proof of Proposition \ref{prop:local-gli} does not yield an
algorithm for computing GLIs, for the following reason. Building an
interpolant for an unsatisfiable conjunction of formulae $\phi \wedge
\psi$ is typically the job of the decision procedure that proves the
unsatisfiability and, in general, there is no such procedure, when
$\phi$ and $\psi$ contain predicates and have non-trivial quantifier
alternation. In this case, some provers use instantiation heuristics
for the universal quantifiers that are sufficient for proving
unsatisfiability, however these heuristics are not always suitable for
interpolant generation. Consequently, from now on, we assume the
existence of an effective Lyndon interpolation procedure only for
decidable theories, such as the quantifier-free linear (integer)
arithmetic with uninterpreted functions (UFLIA, UFLRA, etc.)
\cite{RybalchenkoSofronieStokkermans}.

This is where the predicate-free path formulae (defined in
\S\ref{sec:path-formulae}) come into play. Recall that, for a given
event sequence $\alpha$, the automaton $\A$ accepts a word $w$ such
that $\event{w} = \alpha$ if and only if $\substform{\alpha}$ is
satisfiable (Lemma \ref{lemma:quant-pred-acceptance}). Assuming
further that the equality and interpreted predicates
(e.g.\ inequalities for integers) atoms from the transition rules of
$\A$ belong to a decidable first-order theory, such as Presburger
arithmetic, Lemma \ref{lemma:quant-pred-acceptance} gives us an
effective way of checking emptiness of $\A$, relative to a given event
sequence. However, this method does not cope well with lazy
annotation, because there is no way to extract, from the
unsatisfiability proof of $\substform{\alpha}$, the interpolants
needed to annotate $\alpha$. This is because \begin{inparaenum}[(I)]
\item\label{it1:gli} the formula $\substform{\alpha}$, obtained by
  repeated substitutions loses track of the steps of the execution,
  and
%
\item\label{it2:gli} quantifiers that occur nested in
  $\substform{\alpha}$ make it difficult to write $\substform{\alpha}$
  as an unsatisfiable quantifier-free conjunction of formulae from
  which interpolants are extracted (Definition
  \ref{def:lyndon-interpolant}).
\end{inparaenum}

The solution we adopt for the first issue (\ref{it1:gli}) consists in
partially recovering the time-stamped structure of the acceptance
formula $\accform{\alpha}$ using the formula $\quantform{\alpha}$, in
which only transition quantifiers occur. The second issue
(\ref{it2:gli}) is solved under the additional assuption that the
theory of the data domain $\Data$ has \emph{witness-producing
  quantifier elimination}. More precisely, we assume that, for each
formula $\exists x ~.~ \phi(x)$, there exists an effectively
computable term $\tau$, in which $x$ does not occur, such that
$\exists x ~.~ \phi$ and $\phi[\tau/x]$ are equisatisfiable. These
terms, called \emph{witness terms} in the following, are actual
definitions of the Skolem function symbols from the following folklore
theorem:

\begin{theorem}[\cite{BorgerGraedelGurevich97}]\label{thm:skolem}
  Given $Q_1 x_1 \ldots Q_n x_n ~.~ \phi$ a first-order sentence,
  where $Q_1, \ldots, Q_n \in \set{\exists,\forall}$ and $\phi$ is
  quantifier-free, let $\eta_i \isdef f_i(y_1,\ldots,y_{k_i})$ if $Q_i
  = \forall$ and $\eta_i \isdef x_i$ if $Q_i = \exists$, where $f_i$
  is a fresh function symbol and $\set{y_1, \ldots, y_{k_i}} =
  \set{x_j \mid j < i,~ Q_j = \exists}$. Then the entailment \(Q_1 x_1
  \ldots Q_n x_n ~.~ \phi \models \phi[\eta_1/x_1,\ldots,\eta_n/x_n]\)
  holds.
\end{theorem}
Examples of witness-producing quantifier elimination procedures can be
found in the literature for e.g. linear integer (real) arithmetic
(LIA,LRA), Presburger arithmetic and boolean algebra of sets and
Presburger cardinality constraints (BAPA)
\cite{KuncakMayerPiskacSuter12}.

Under the assumption that witness terms can be effectively built, we
describe the generation of a non-local GLI for a given input event
sequence $\alpha = a_1 \ldots a_n$. First, we generate successively
the acceptance formula $\accform{\alpha}$ and its equisatisfiable
forms $\quantform{\alpha} = Q_1x_1 \ldots Q_mx_m ~.~ \widehat{\Phi}$
and $\substform{\alpha} = Q_1x_1 \ldots Q_mx_m ~.~ \overline{\Phi}$,
both written in prenex form, with matrices $\widehat{\Phi}$ and
$\overline{\Phi}$, respectively. Because we assumed that the first
order theory of $\Data$ has quantifier elimination, the satisfiability
problem for $\substform{\alpha}$ is decidable. If $\substform{\alpha}$
is satisfiable, we build a counterexample for emptiness $w$ such that
$\event{w}=\alpha$ and $\data{w}$ is a satisfying assignment for
$\substform{\alpha}$. Otherwise, $\substform{\alpha}$ is unsatisfiable
and there exist witness terms $\tau_{i_1} \ldots \tau_{i_\ell}$, where
$\set{i_1, \ldots, i_\ell} = \set{j \in [1,m] \mid Q_j = \forall}$,
such that $\overline{\Phi}[\tau_{i_1}/x_{i_1}, \ldots,
  \tau_{i_\ell}/x_{i_\ell}]$ is unsatisfiable (Theorem
\ref{thm:skolem}). Then it turns out that the formula
$\widehat{\Phi}[\tau_{i_1}/x_{i_1}, \ldots,
  \tau_{i_\ell}/x_{i_\ell}]$, obtained analogously from the matrix of
$\quantform{\alpha}$, is unsatisfiable as well (Lemma
\ref{lemma:subst-quant} below). Because this latter formula is
structured as a conjunction of formulae $\stamp{\iota}{0} \wedge
\phi_1 \ldots \wedge \phi_n \wedge \psi$, where $\voc{\phi_k} \cap
\stamp{Q}{\leq n} \subseteq \stamp{Q}{k-1} \cup \stamp{Q}{k}$ and
$\voc{\psi} \cap \stamp{Q}{\leq n} \subseteq \stamp{Q}{n}$, it is now
possible to use an existing interpolation procedure for the
quantifier-free theory of $\Data$, extended with uninterpreted
function symbols, to compute a (not necessarily local) GLI $(I_0,
\ldots, I_n)$ such that $\voc{I_k} \cap \stamp{Q}{\leq n} \subseteq
\stamp{Q}{k}$, for all $k \in [n]$.

\begin{example}[Contd. from Examples \ref{ex:quant-elim} and
    \ref{ex:pred-elim}]\label{ex:gli} The formula $\substform{\alpha}$
  (Example \ref{ex:pred-elim}) is unsatisfiable and let $\tau_2 \isdef
  z_1$ be the witness term for the universally quantified variable
  $z_2$. Replacing $z_2$ with $\tau_2$ ($z_1$) in the matrix of
  $\quantform{\alpha}$ (Example \ref{ex:quant-elim}) yields the
  unsatisfiable conjunction below, obtained after trivial
  simplifications:
\[\begin{array}{lcll}
           [z_1 \geq 0 \wedge \stamp{q}{0}(z_1)] & \wedge &
           [\stamp{q}{0}(z_1) \rightarrow \stamp{x}{1} \geq 0 \wedge \stamp{q}{1}(\stamp{x}{1}+z_1)] ~\wedge \\
           && [\stamp{q}{1}(\stamp{x}{1}+z_1) \rightarrow \stamp{x}{1}+z_1 < 0 \wedge \stamp{q_f}{2}(\stamp{x}{2}+\stamp{x}{1}+z_1)]
\end{array}\]
A non-local GLI for the above conjunction is the sequence of formulae: 
\[(\stamp{q}{0}(z_1) \wedge z_1\geq 0,~ \stamp{x}{1} \geq 0 \wedge \stamp{q}{1}(\stamp{x}{1}+z_1)
\wedge z_1\geq0,~\bot) \hspace*{4cm}\blacksquare\]
\end{example}

We formalize and prove the correctness for the above construction of
non-local GLI. A function $\xi : \nat \rightarrow \nat$ is
\emph{monotonic} iff for each $n < m$ we have $\xi(n) \leq \xi(m)$ and
\emph{finite-range} iff for each $n \in \nat$ the set $\set{m \mid
  \xi(m)=n}$ is finite. If $\xi$ is finite-range, we denote by
$\xi_{\max}^{-1}(n) \in \nat$ the maximal value $m$ such that
$\xi(m)=n$.
\begin{lemma}\label{lemma:subst-quant}
  Given a non-empty input event sequence $\alpha = a_1 \ldots a_n \in
  \Sigma^*$, such that $\accform{\alpha}$ is unsatisfiable, let
  $Q_1x_1 \ldots Q_mx_m ~.~ \widehat{\Phi}$ be a prenex form of
  $\quantform{\alpha}$ and let $\xi : [1,m] \rightarrow [n]$ be a
  monotonic finite-range function mapping each transition quantifier
  to the minimal index from the sequence $\quantformn{\alpha_0},
  \ldots, \quantformn{\alpha_n}$ where it occurs. Then one can
  effectively build:
  \begin{compactenum}
  \item\label{it1:subst-quant} witness terms $\tau_{i_1}, \ldots,
    \tau_{i_\ell}$, where \(\set{i_1, \ldots,i_\ell} = \set{j \in
      [1,m] \mid Q_j = \forall}\) and \(\voc{\tau_{i_j}} \subseteq
    \stamp{X}{\leq \xi(i_j)} \cup \set{x_k \mid k < i_j, Q_k =
      \exists},~ \forall j \in [1,\ell]\) such that
    $\widehat{\Phi}[\tau_{i_1}/x_{i_1}, \ldots,
      \tau_{i_\ell}/x_{i_\ell}]$ is unsatisfiable, and
    %
  \item\label{it2:subst-quant} a GLI $(I_0, \ldots, I_n)$ for
    $\alpha$, such that \(\voc{I_k} \subseteq \stamp{Q}{k} \cup
    \stamp{X}{\leq k} \cup \set{x_j \mid j < \xi_{\max}^{-1}(k),~ Q_j
      = \exists}\), for all $k \in [n]$.
  \end{compactenum}
\end{lemma}

Consequently, under two assumptions about the first-order theory of
the data domain, namely\ \begin{inparaenum}[(i)]
\item witness-producing quantifier elimination, and
\item Lyndon interpolation for the quantifier-free fragment with
  uninterpreted functions,
\end{inparaenum}
we developed a generic method that produces GLIs for unfeasible input
event sequences. Moreover, each formula in the interpolant refers only
to the current predicate symbols, the current and past input variables
and the existentially quantified transition variables introduced at
the previous steps. The remaining questions are how to use these GLIs
to label the sequences in the unfolding of an automaton (Definition
\ref{def:unfolding}) and compute coverage (Definition
\ref{def:coverage}) between nodes of the unfolding.

\subsection{Unfolding with Non-local Interpolants}

As required by Definition \ref{def:unfolding}, the unfolding $U$ of an
automaton $\A = \tuple{\Sigma,X,Q,\iota,F,\Delta}$ is labeled by
formulae $U(\alpha) \in \posforms(Q,\emptyset)$, with no free symbols,
other than predicate symbols, such that the labeling is compatible
with the transition relation of the automaton. Each newly expanded
input sequence of $\A$ is initially labeled with $\top$ and the labels
are refined using GLIs computed from proofs of spuriousness. The
following lemma describes the refinement of the labeling of an input
sequence by a non-local GLI:

\begin{lemma}\label{lemma:refinement}
  Let $U$ be an unfolding of an automaton $\A =
  \tuple{\Sigma,X,Q,\iota,F,\Delta}$ such that $\alpha = a_1\ldots a_n
  \in \dom(U)$ and $(I_0, \ldots, I_n)$ is a GLI for $\alpha$. Then
  the mapping $U' : \dom(U) \rightarrow \posforms(Q,\emptyset)$ is an
  unfolding of $\A$, where:
  \begin{compactitem}
  \item $U'(\alpha_k) = U(\alpha_k) \wedge J_k$, for all $k \in [n]$,
    where $J_k$ is the formula obtained from $I_k$ by removing the
    time stamp of each predicate symbol $\stamp{q}{k}$ and
    existentially quantifying each free variable, and
  %
\item $U'(\beta) = U(\beta)$ if $\beta \in \dom(U)$ and $\beta
  \not\prefix \alpha$,
  \end{compactitem}
  Moreover, $\alpha$ is safe in $U'$.
\end{lemma}
Observe that, by Lemma \ref{lemma:subst-quant}
(\ref{it2:subst-quant}), the set of free variables of a GLI formula
$I_k$ consists of \begin{inparaenum}[(i)]
\item variables $\stamp{X}{\leq k}$ keeping track of data values seen
  in the input at some earlier moment in time, and
%
\item variables that track past choices made within the transition
  rules.
\end{inparaenum}
Basically, it is not important when exactly in the past a certain
input has been read or when a choice has been made, because only the
relation between the values of these and the current variables
determines the future behavior of the automaton. Quantifying these
variables existentially does the job of ignoring when exactly in the
past these values have been seen. Moreover, the last point of Lemma
\ref{lemma:refinement} ensures that the refined path is safe in the
new unfolding and will stay safe in all future refinements of this
unfolding.

The last ingredient of the lazy annotation semi-algorithm based on
unfoldings consist in the implementation of the coverage check, when
the unfolding of an automaton is labeled with conjunctions of
existentially quantified formulae with predicate symbols, obtained
from interpolation. By Definition \ref{def:coverage}, checking whether
a given node $\alpha \in \dom(U)$ is covered amounts to finding a
prefix $\alpha' \prefix \alpha$ and a node $\beta \in \dom(U)$ such
that $U(\alpha') \models U(\beta)$, or equivalently, the formula
$U(\alpha') \wedge \neg U(\beta)$ is unsatisfiable. However, the
latter formula, in prenex form, has quantifier prefix in the language
$\exists^*\forall^*$ and, as previously mentioned, the satisfiability
problem for such formulae becomes undecidable when the data theory
subsumes Presburger arithmetic \cite{Halpern91}.

Nevertheless, if we require just a yes/no answer (i.e.\ not an
interpolant) recently developed quantifier instantiation heuristics
\cite{ReynoldsKK17} perform rather well in answering a large number of
queries in this class. Observe, moreover, that coverage does not need
to rely on a complete decision procedure. If the prover fails in
answering the above satisfiability query, then the semi-algorithm
assumes that the node is not covered and continues exploring its
successors. Failure to compute complete coverage may lead to
divergence (non-termination) and ultimately, to failure to prove
emptiness, but does not affect the soundness of the semi-algorithm
(real counterexamples will still be found).

\section{Experimental Results}

We have implemented a version of the IMPACT semi-algorithm
\cite{McMillan06} in a prototype tool, avaliable
online \cite{foada}. The tool is
written in Java and uses the Z3 SMT solver \cite{z3}, via the JavaSMT
interface \cite{javasmt}, for spuriousness and coverage queries and
also for interpolant generation. Table \ref{tab:experiments} reports
the size of the input automaton in bytes, the numbers of Predicates,
Variables and Transitions, the result of emptiness check, the number
of Expanded and Visited Nodes during the unfolding and the Time in
miliseconds.  The experiments were carried out on a MacOS x64 - 1.3
GHz Intel Core i5 - 8 GB 1867 MHz LPDDR3 machine.

\begin{table}[htb]
\vspace*{-\baselineskip}
\begin{center}
{\fontsize{7}{8}\selectfont
\begin{tabular}{||l|c|c|c|c|c|c|c|c||}
\hline
Example & $\len{\A}$ (bytes) & Predicates & Variables & Transitions & $L(\A)=\emptyset$ ? & Nodes Expanded & Nodes Visited & Time (msec) \\
\hline
incdec.pa & 499 & 3 & 1 & 12 & no & 21 & 17 & 779 \\
\hline
localdec.pa & 678 & 4 & 1 & 16 & no & 49 & 35 & 1814 \\
\hline
ticket.pa & 4250 & 13 & 1 & 73 & no & 229 & 91 & 9543 \\
\hline
count\_thread0.pa & 9767 & 14 & 1 & 126 & no & 154 & 128 & 8553 \\
\hline
count\_thread1.pa & 10925 & 15 & 1 & 135 & no & 766 & 692 & 76771 \\
\hline
local0.pa & 10595 & 13 & 1 & 117 & no & 73 & 27 & 1431 \\
\hline
local1.pa & 11385 & 14 & 1 & 126 & no & 1135 & 858 & 101042 \\
\hline
array\_rotation.ada & 1834 & 8 & 7 & 7 & yes & 9 & 8 & 1543 \\
\hline
array\_simple.ada & 3440 & 9 & 5 & 8 & yes & 11 & 10 & 6787 \\
\hline
array\_shift.ada & 874 & 6 & 5 & 5 & yes & 6 & 5 & 413 \\
\hline
abp.ada & 6909 & 16 & 14 & 28 & no & 52 & 47 & 4788 \\
\hline
train.ada & 1823 & 10 & 4 & 26 & yes & 68 & 67 & 7319 \\
\hline
hw1.ada & 322 & 3 & 2 & 5 & Solver Error & / & / & / \\
\hline
hw2.ada & 674 & 7 & 2 & 8 & yes & 20 & 22 & 4974 \\
\hline
rr-crossing.foada & 1780 & 10 & 1 & 16 & yes & 67 & 67 & 7574 \\
% \hline
% running2.foada & 1217 & yes & 13 & 13 & 788 \\
\hline
train-simple1.foada & 5421 & 13 & 1 & 61 & yes & 43 & 44 & 2893 \\
\hline
train-simple2.foada & 10177 & 16 & 1 & 118 & yes & 111 & 113 & 8386 \\
\hline
train-simple3.foada & 15961 & 19 & 1 & 193 & yes & 196 & 200 & 15041 \\
% \hline
% fischer-mutex1.foada & 1566 & no & 31 & 31 & 1552 \\
\hline
fischer-mutex2.foada & 3000 & 11 & 2 & 23 & yes & 23 & 23 & 808 \\
\hline
fischer-mutex3.foada & 4452 & 16 & 2 & 34 & yes & 33 & 33 & 1154 \\
\hline
\end{tabular}
}
\caption{Experiments with First Order Alternating Automata}
\label{tab:experiments}
\end{center}
\vspace*{-3\baselineskip}
\end{table}

The test cases shown in Table \ref{tab:experiments}, come from several
sources, namely predicate automata models (*.pa)
\cite{Farzan15,Farzan16} available online \cite{pa}, timed automata
inclusion problems ({\tt abp.ada}, {\tt train.ada}, {\tt
  rr-crossing.foada}), array logic entailments ({\tt array\_rota-}
{\tt tion.ada}, {\tt array\_simple.ada}, {\tt array\_shift.ada}) and
hardware circuit verification ({\tt hw1.ada}, {\tt hw2.ada}),
initially considered in \cite{IosifRV16}, with the restriction that
local variables are made visible in the input. The {\tt
  train-simpleN.} {\tt foada} and {\tt fischer-mutexN.} {\tt foada}
examples are parametric verification problems in which one checks
inclusions of the form $\bigcap_{i=1}^N\lang{A_i} \subseteq \lang{B}$,
where $A_i$ is the $i$-th copy of the template automaton.

The advantage of using FOADA over the INCLUDER \cite{includer} tool
from \cite{IosifRV16} is the possibility of having automata over
infinite alphabets with local variables, whose values are not visible
in the input. In particular, this is essential for checking inclusion
of timed automata that use internal clocks to control the computation.

\section{Conclusions}

We present first-order alternating automata, a model of computation
that generalizes classical boolean alternating automata to first-order
theories. Due to their expressivity, first-order alternating automata
are closed under union, intersection and complement. However the
emptiness problem is undecidable even in the most simple case, of the
quantifier-free theory of equality with uninterpreted predicate
symbols. We deal with the emptiness problem by developping a practical
semi-algorithm that always terminates, when the automaton is not
empty. In case of emptiness, termination of the semi-algorithm occurs
in most practical test cases, as shown by a number of experiments.