### 2019-07-04 quantifiers/rebuttal.txt
ok

We first wish to thank all reviewers for their time and great
work. Below we answer the concerns raised by Reviewers 1 and 4.

Review 1:

 1.1 Relation of FOAA with s-AFA.

 Symbolic (alternating) finite-state automata are an important
 extension of finite-state automata to infinite alphabets and lacking
 those references in our paper is a minus that should be
 corrected. The reason is our original interest in language
 recognizers relating values occurring at different positions in the
 input, by means of internal variables: clocks for timed automata (TA)
 and registers for finite-memory automata [13]. Our motivation is not
 as much finding decidable infinite alphabet models but rather having
 a very general model, closed under complement, union and intersection
 (boolean closure is merely a consequence of its expressive power), in
 which all undecidable inclusion problems can be encoded, then devise
 a practical semi-algorithm for it.

 Although the finite event alphabet \Sigma can be encoded in the input
 data (it is introduced for technical convenience), modeling control
 by uninterpreted predicates with internal variables as arguments is
 key to capturing computations using hidden variables (pg 2, first
 paragraph). Note that Definition 1 (point 2) requires synchronization
 of all execution branches on all input events and values.

 To the best of our knowledge, there are no published semi-algorithms
 for TA or RA inclusion (tools such as UPAAL check only for
 emptiness). Our method contributes to these areas and to program
 verification (Section 5 considers several programs handling integer
 arrays).

 1.2 Restrictions of FOAA with decidable emptiness problems.

 As proved in [5], emptiness of predicate automata is undecidable even
 for quantifier-free rules with equalities and disequalities, not
 using the input variables, when predicates have arities of two or
 more (having only equalities and monadic predicates yields a
 decidable model though). In the light of this result (pg. 9, last
 paragraph), any reasonably interesting extension of monadic predicate
 automata is likely to have an undecidable emptiness problem.

 We also compare with an alternating automaton model [11], in which
 states are propositional variables (no uninterpreted predicates) and
 comparisons are possible only between adjacent input values, without
 hidden variables. This model is intermediate between s-AFA and FOAA,
 nevertheless emptiness remains undecidable.

Review 4:

4.1 Relation with Constrained Horn Clauses (CHC).

 The main difference between the unfoldings of CHC and those of FOAA
 is that, in CHC, the different branches of the unfolding tree are
 unrelated (at arbitrary depths), whereas in FOAA, all branches of the
 computation tree must synchronize on the same input word. As
 suggested, to reflect this synchronization, a CHC encoding of FOADA
 must use data lists to ensure that, at each step, all incoming
 branches consider exactly the same list, encoding the continuation of
 the input from that point on (CHC traverse the input from the end to
 the beginning). An encoding of FOAA in CHC would benefit from the
 advantage of using off-the-shelf CHC solvers, but, in our opinion,
 would face several problems, listed below in increasing order of
 difficulty:

I. Checking that, in each clause, all incoming branches read the same
 input can be done using equalities between the list arguments of the
 subgoal predicates. However, the body of the clause must be given
 either as a conjunction or, if quantifiers are considered, as prenex
 DNF. However, turning each rule into DNF would potentially add to the
 complexity.

II. Introducing recursive algebraic data types (ADT) in the encoding
 requires an interpolation procedure for ADTs in the emptiness
 semi-algorithm. Although satisfiability of quantifier-free ADTs is
 mature and integrated with many SMT solvers, interpolation of ADT
 theories is at its beginnings. In fact, the technical report [*]
 claims to introduce the first interpolation procedure for ADTs:

 * Hossein Hojjat, Philipp Ruemmer. Deciding and Interpolating
 Algebraic Data Types by Reduction. ArXiv:1801.02367. Proc.
 SYNASC'17.

 Moreover, the availability of such procedures in off-the-shelf CHC
 solvers is still an unresolved issue. For instance, the integration
 of the above procedure in ELDARICA is underway, according to
 [*]. However, when such technical problems will be solved, it would
 definitely be interesting to have an experimental comparison between
 FOAA and ELDARICA (running on an encoding of our examples using
 ADTs).

III. Having quantified rules in FOADA requires a quantifier
 elimination procedure that produces witness terms. However, combining
 quantifiers and ADTs results most likely in an undecidable theory,
 for which no quantifier elimination exists (by an argument similar to
 the one of [8]).

 Essentially, for the above reasons, we have chosen to design our
 emptiness checking semi-algorithm from scratch, imposing as few
 assumptions as possible on the underlying first-order data theory. In
 fact, we only assume (a) Lyndon interpolation for quantifier-free
 formulae with uninterpreted predicates (without ADTs) and (b) witness
 term generation during quantifier elimination. Both assumptions are
 fairly reasonable for standard SMT theories (e.g. UFLIA, UFLRA).